### 2020-02-27 drugspending.Rmd
medicare drug spending cleaned

---
title: "Medicare Part B Drug Spending"
author: "Suruchi Ramanujan"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(janitor)
```

```{r}
drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
clean_names() %>%
  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units",
         ,"x2014_total_spending", "x2014_total_dosage_units", ,"x2015_total_spending", "x2015_total_dosage_units", ,"x2016_total_spending", "x2016_total_dosage_units", ,"x2017_total_spending", "x2017_total_dosage_units",
         "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
```

### 2020-02-28 drugspending.Rmd
drug spending and opioid death distribution

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 library(dplyr)
@@
 ```{r}
+
+# This data tracks the spending over a course of five years of a series of
+# opioids. Perhaps, we will find a correlation between spending and drug type
+# for opioid deaths.
+
 drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
 clean_names() %>%
-  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units",
-         ,"x2014_total_spending", "x2014_total_dosage_units", ,"x2015_total_spending", "x2015_total_dosage_units", ,"x2016_total_spending", "x2016_total_dosage_units", ,"x2017_total_spending", "x2017_total_dosage_units",
+  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", "x2015_total_spending", "x2015_total_dosage_units",
+  "x2016_total_spending", "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
          "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
 ```
+
+```{r}
+
+# This data set displays the number of deaths per county in NY over a course of
+# 15 years.
+
+deathbycounty_ny<- read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
+  clean_names()
+```
+
+Introduction: This project is meant to look into trends regarding the issue of opioid overuse.
+The misuse of opioids has ignited a national crisis that has contributed to worsening health outcomes for thousands and their families throughout the United States.
+This
+
+Data set 1:
+
+Data set 2:
+
+Data set 3:
+
+Data source: https://about.twitter.com/en_us/values/elections-integrity.html#data
+
+About Me: My name is Suruchi Ramanujan and I am a senior in Quincy House studying Molecular and Cellular Biology.
+I grew up in Overland Park, Kansas so this year has been a real win for me (Go Chiefs).
+In my free time, I enjoy performing Indian Classical dance, running, and grinding through my DataCamp exercises.
+
+My github repo for this project is below:

### 2020-02-28 opioidcrisis.Rmd
Need to add a little more to intro then done?

---
title: "Medicare Part B Drug Spending"
author: "Suruchi Ramanujan"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(janitor)
```

```{r}

# This data tracks the spending over a course of five years of a series of
# drugs. I want to compare the spending levels of the different kinds of drugs
# by categorizing them based on their use. Perhaps, we will find a correlation
# between spending and drug type for opioid deaths.

drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
clean_names() %>%
  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", "x2015_total_spending", "x2015_total_dosage_units",
  "x2016_total_spending", "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
         "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
```

```{r}

# This data set displays the number of deaths caused by opioids per county in NY
# over a course of 15 years.

deathbycounty_ny<- read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
  clean_names()

# Using data commons, I isolated drug spending in NY on some of the worst drugs
# of addiction and treatment drugs such as Methadone, Buprenorphine, and Naloxone.

drugspending_ny<- read.csv("New York Drug Spending.csv") %>%
  clean_names() %>%
  rename( Buprenorphine = "new_york_drug_dea_9064") %>%
  rename( Cocaine = "new_york_drug_dea_9041l") %>%
  rename( Methadone = "new_york_drug_dea_9250b") %>%
  rename( Naloxone = "new_york_drug_dea_9411")

# I want to compare the drug spending in NY (a state that has a lot of opioid
# data with the country that has the worst opioid addiction rates, West
# Virginia)

drugspending_wv<- read.csv("West Virginia Drug Spending.csv") %>%
clean_names() %>%
  rename( Buprenorphine = "west_virginia_drug_dea_9064") %>%
  rename( Cocaine = "west_virginia_drug_dea_9041l") %>%
  rename( Methadone = "west_virginia_drug_dea_9250b") %>%
  rename( Naloxone = "west_virginia_drug_dea_9411")

```

Introduction: This project is meant to look into trends regarding the issue of opioid overuse. The misuse of opioids has ignited a national crisis that has contributed to worsening health outcomes for thousands and their families throughout the United States. This

Data set 1: Data Set Tracks Overall trends in drug spending amongst opioids and nonopioids across the country. I want to look at this just to see which types of drugs are receiving the most attention and is it with reason that these drugs are being spent on.

Data set 2: Data Set 2 introduces the audience to opioid deaths by county in New York from 2003- 2017. I imagine using a heat map to see how levels have changed over time/ whether there is a trend in opioid deaths.

Data set 3: Data Set 3 outlines the spending in New York on one of the worst addiction drugs (Cocaine) vs. three of the most widely used addiction treatment drugs (Methadone, Naloxone, and Buprenorphine). This site has information for other drugs too so I may include those as well. I still have to decide which drugs to include though.

Data Set 4: West Virginia is one of the worst hit states in terms of opioid addiction. I would like to compare the drug spending for the same four drugs in Data Set 3 between West Virginia and New York and note any trends/comparisons.

Data sources:For Data sets 1 and 2, I used healthdata.gov. For Data sets 3 and 4, I used Data Commons.

About Me: My name is Suruchi Ramanujan and I am a senior in Quincy House studying Molecular and Cellular Biology. I grew up in Overland Park, Kansas so this year has been a real win for me (Go Chiefs). In my free time, I enjoy performing Indian Classical dance, running, and grinding through my DataCamp exercises.

My github repo for this project is below: https://github.com/suruchiramanujan/final-project.git

### 2020-02-28 opioidcrisis.Rmd
ms3 complete

@@
 Introduction: This project is meant to look into trends regarding the issue of opioid overuse.
-The misuse of opioids has ignited a national crisis that has contributed to worsening health outcomes for thousands and their families throughout the United States.
-This
+In the 1990's doctors overprescribed pain relievers as a temporary bandage to disease without knowing the potential for addiction.
+Since then, the misuse of opioids has ignited a national crisis that has contributed to worsening health outcomes for thousands and their families throughout the United States.
+Opioid misuse is not only an emotional burden for patients and their families, it also loses the United States billions of dollars yearly.
+Thus, this study looks into some of the spending trends of opioid use in conjunction with death rates.
 

### 2020-03-07 ms_4.Rmd
drug deaths over time

--- 2020-02-28 opioidcrisis.Rmd
@@
 ---
-title: "Medicare Part B Drug Spending"
+title: "ms_4"
 author: "Suruchi Ramanujan"
-date: "2/27/2020"
+date: "3/6/2020"
 output: html_document
 ---
+---
+title: "ms_4"
+author: "Suruchi Ramanujan"
+date: "3/6/2020"
+output: html_document
+---
+** This is a modified version of ms_3 since I answered many of the questions for ms_4 in last week's analysis.*
 
@@
 
+These first ___ are the new data sets
 ```{r}
+# This data provides state-specific numbers regarding the number of drug
+# overdose deaths per state as well as the predicted numbers for drug overdose
+# deaths. I am super excited to plot this graph onto a map and look at how the
+# worst affected states change over time. I am also interested in seeing whether
+# trends match across drug groups.
 
+drugoverdosedeaths<- read.csv("VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
+  clean_names() %>%
+  select("state", "year", "month", "indicator", "data_value", "state_name", "predicted_value")
+```
+
+```{r}
 # This data tracks the spending over a course of five years of a series of
@@
 clean_names() %>%
-  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", "x2015_total_spending", "x2015_total_dosage_units",
-  "x2016_total_spending", "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
+  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", 
+  "x2015_total_spending", "x2015_total_dosage_units", "x2016_total_spending", 
+  "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
   "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
@@
 
-deathbycounty_ny<- read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
+deathbycounty_ny<- 
+  read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
   clean_names()
@@
 
-Data set 1: Data Set Tracks Overall trends in drug spending amongst opioids and nonopioids across the country.
+Data Set 1: Data Set Tracks Overall trends in drug spending amongst opioids and nonopioids across the country.
 I want to look at this just to see which types of drugs are receiving the most attention and is it with reason that these drugs are being spent on.
 
-Data set 2: Data Set 2 introduces the audience to opioid deaths by county in New York from 2003- 2017.
+Data Set 2: Data Set 2 introduces the audience to opioid deaths by county in New York from 2003- 2017.
 I imagine using a heat map to see how levels have changed over time/ whether there is a trend in opioid deaths.
 
-Data set 3: Data Set 3 outlines the spending in New York on one of the worst addiction drugs (Cocaine) vs.
+Data Set 3: Data Set 3 outlines the spending in New York on one of the worst addiction drugs (Cocaine) vs.
 three of the most widely used addiction treatment drugs (Methadone, Naloxone, and Buprenorphine).
@@
 
-Data sources:For Data sets 1 and 2, I used healthdata.gov. For Data sets 3 and 4, I used Data Commons.
+Data sources:For Data Sets 1 and 2, I used healthdata.gov. For Data Sets 3 and 4, I used Data Commons.
+
+Future: I think I am content with the state specific data with drugs of choice but I think it would be cool to also map out opioid deaths based on country region or based on income levels in a location.
+I still have to flesh out this idea though.
 

### 2020-03-07 ms_4.Rmd
final commit?

--- 2020-02-28 opioidcrisis.Rmd
@@
 ---
-title: "Medicare Part B Drug Spending"
+title: "ms_4"
 author: "Suruchi Ramanujan"
-date: "2/27/2020"
+date: "3/6/2020"
 output: html_document
 ---
+---
+title: "ms_4"
+author: "Suruchi Ramanujan"
+date: "3/6/2020"
+output: html_document
+---
+** This is a modified version of ms_3 since I answered many of the questions for ms_4 in last week's analysis.**
 
@@
 library(janitor)
+library(tidyverse)
 ```
@@
 
+# This first data set is the new data set for the week. This data provides
+# state-specific numbers from the CDC regarding the number of drug overdose
+# deaths per state as well as the predicted numbers for drug overdose deaths. I
+# am super excited to plot this graph onto a map and look at how the worst
+# affected states change over time.
+
+drugoverdosedeaths<- read.csv(
+  "VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
+  clean_names() %>%
+  select("state", "year", "month", "indicator", 
+         "data_value", "state_name", "predicted_value") %>%
+
+# I only want the information rows containing drug death information- I still
+# need to decipher the difference between "Deaths" and "Drug Overdose Deaths"
+# though.
+
+ filter(str_detect(indicator, "Death")) %>%
+
+# Ultimately, I want to look at state-specific data/trends so I want to group by
+# state.
+
+ group_by("state_name")
+```
+
+```{r}
 # This data tracks the spending over a course of five years of a series of
@@
 
-Data set 1: Data Set Tracks Overall trends in drug spending amongst opioids and nonopioids across the country.
+Data Set 1: Data Set 1, from the CDC, introduces drug overdose death levels over the last couple of years based on state.
+Visually, for readers, I think this is very convenient data for recognizing which states are worst hit by the opioid crisis.
+Just to briefly look into which states are worst hit, I did an arrange(desc()) and saw that the data agreed with website information online that said West Virginia is worst hit by the crisis; however, I also noticed that some data is incorrectly coded/is missing so I will have to figure out how to work with that.
+It looks like West Virginia, South Carolina, and Colorado are worst hit by the epidemic so I might add a couple more Data Sets using Data Commons that analyze opioid spending in S.
+Carolina and Colorado in comparison to the existing Data Set 5.
+It might also be interesting to look at which state is least hit by the epidemic.
+Writing this, however, I think it might be important to normalize the death counts to state population as well, so I might join a table with state populations for the years indicated as well.
+
+Data Set 2: Data Set tracks overall trends in drug spending amongst opioids and nonopioids across the country.
 I want to look at this just to see which types of drugs are receiving the most attention and is it with reason that these drugs are being spent on.
 
-Data set 2: Data Set 2 introduces the audience to opioid deaths by county in New York from 2003- 2017.
+Data Set 3: Data Set 2 introduces the audience to opioid deaths by county in New York from 2003- 2017.
 I imagine using a heat map to see how levels have changed over time/ whether there is a trend in opioid deaths.
 
-Data set 3: Data Set 3 outlines the spending in New York on one of the worst addiction drugs (Cocaine) vs.
+Data Set 4: Data Set 3 outlines the spending in New York on one of the worst addiction drugs (Cocaine) vs.
 three of the most widely used addiction treatment drugs (Methadone, Naloxone, and Buprenorphine).
@@
 
-Data Set 4: West Virginia is one of the worst hit states in terms of opioid addiction.
+Data Set 5: West Virginia is one of the worst hit states in terms of opioid addiction.
 I would like to compare the drug spending for the same four drugs in Data Set 3 between West Virginia and New York and note any trends/comparisons.
+I imagine comparative line graphs but am definitely open to changing the visual model based on the different visualization methods I learn in class.
 
-Data sources:For Data sets 1 and 2, I used healthdata.gov. For Data sets 3 and 4, I used Data Commons.
+Data sources:For Data Set 1, I used information from the CDC. For Data Sets 2 and 3, I used healthdata.gov. For Data Sets 4 and 5, I used Data Commons.
+
+Future: I found this site: https://pdas.samhsa.gov/#/ with extremely helpful data sources, but I am not sure which categories I would like to compare given how much data there is available.
+I think by next week I would like to further narrow my interests in opioid addiction and decide if I even want to use this data.
 

### 2020-03-20 ms_5.Rmd
new ggplots!!

---
title: "ms_5"
author: "Suruchi Ramanujan"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(janitor)
library(tidyverse)
```

*Copy pasted from last week
```{r}

# This first data set is the new data set from last week. This data provides
# state-specific numbers from the CDC regarding the number of drug overdose
# deaths per state as well as the predicted numbers for drug overdose deaths. I
# am super excited to plot this graph onto a map and look at how the worst
# affected states change over time.

drugoverdosedeaths<- read.csv(
  "VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
  clean_names() %>%
  select("state", "year", "month", "indicator", 
         "data_value", "state_name", "predicted_value") %>%

# I only want the information rows containing drug death information- I still
# need to decipher the difference between "Deaths" and "Drug Overdose Deaths"
# though.

 filter(str_detect(indicator, "Death")) %>%

# Ultimately, I want to look at state-specific data/trends so I want to group by
# state.

 group_by("state_name")
```

*New for this week
```{r}
# For this week, I am going to plot the Drug Overdose Deaths in April of 2015 by
# state according to the CDC. I want to use this data just because it is the
# first recorded date of drug overdose deaths. I am hoping there will be some
# expected trend across states.

drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)

drugodapr15<- drugoverdosedeaths %>%
  arrange(desc(data_value)) %>%
  select(state_name, `"state_name"`, data_value, indicator, month, year) %>%
  filter(year == "2015",
        month == "April",
        indicator == "Number of Drug Overdose Deaths")

ggplot(drugodapr15, aes(x= reorder(state_name,-data_value), y= data_value)) +
geom_col() +
labs(title = "Drug Overdose Deaths by State (April 2015)", 
     x = "State", 
     y = "Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

# I am going to do the same with July 2019, which is the latest date of
# recording.

drugodjul19<- drugoverdosedeaths %>%
  arrange(desc(data_value)) %>%
  select(state_name, `"state_name"`, data_value, indicator, month, year) %>%
  filter(year == "2019",
        month == "July",
        indicator == "Number of Drug Overdose Deaths")

ggplot(drugodjul19, aes(x= reorder(state_name,-data_value), y= data_value)) +
geom_col() +
labs(title = "Drug Overdose Deaths by State (July 2019)", 
     x = "State", 
     y = "Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
difference_over_time<- drugodapr15 %>%
  full_join(drugodjul19, by = "state_name") %>%
  mutate(difference = data_value.y - data_value.x)

ggplot(difference_over_time, aes(x= reorder(state_name,-difference), y= difference)) +
geom_col() +
labs(title = "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
     x = "State", 
     y = "Difference in Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 2020-03-20 ms_5.Rmd
ms 5 complete!

@@
 knitr::opts_chunk$set(echo = FALSE)
+
+# I want to load all necessary packages.
+
 library(dplyr)
@@
 
-*Copy pasted from last week
-```{r}
+*Reading in drug overdose data was copy pasted from last week
+```{r, echo = FALSE}
 
@@
 
-*New for this week
-```{r}
+*The graphs below are new for this week. I don't know that I am satisfied with the readability of the last graph and will probably change it slightly in my final version.
+
+```{r, echo = FALSE}
+
 # For this week, I am going to plot the Drug Overdose Deaths in April of 2015 by
@@
 
+# I am not sure if this makes a difference but I am going to change the variable
+# to be numeric (from factor) to rank later on.
+
 drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
+
+# Here, I want to rank the deaths by state in descending order and filter them
+# such that we are only looking at info from April of 2015.
 
@@
 
+# Now, I want to ggplot the data. In the aesthetics, when I simply plot the
+# state name vs data value, the data values do not rank in descending order
+# despite specifying earlier, so I used the trick from PSet 3 of reordering in
+# the x specifications.
+
 ggplot(drugodapr15, aes(x= reorder(state_name,-data_value), y= data_value)) +
 geom_col() +
+
+# I want to label the title and x and y axes so that they are more descriptive.
+
 labs(title = "Drug Overdose Deaths by State (April 2015)", 
@@
      y = "Drug Overdose Deaths") +
+
+# I changed the theme such that the backing of the graph is blank.
+
 theme_classic()+
+
+# The x axis details are cluttered so I chose to angle them for easy reading.
+
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 # I am going to do the same with July 2019, which is the latest date of
-# recording.
+# recording. All code is the same except that the filter specifications at the
+# start are for July of 2019 and I changed the title to specify July of 2019.
 
@@
 
-```{r}
+```{r, echo = FALSE}
+
+# I now want to look at how drug overdose death levels changed between 2019 and
+# 2015 by state. To do so, I combined the two data tables from above using a
+# full join.
+
 difference_over_time<- drugodapr15 %>%
   full_join(drugodjul19, by = "state_name") %>%
+
+# I also want to add a column which indicates the difference in deaths between
+# the two data tables.
+
 mutate(difference = data_value.y - data_value.x)
 
-ggplot(difference_over_time, aes(x= reorder(state_name,-difference), y= difference)) +
+# I want to graph the difference over time by state with the x axis ranking the
+# difference decrease.
+
+ggplot(difference_over_time, aes(x= reorder(state_name,-difference), 
+                                 y= difference)) +
 geom_col() +
-labs(title = "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
+
+# I also changed the title label and the y axis label to specify the
+# differences.
+
+labs(title = 
+    "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
      x = "State", 
      y = "Difference in Drug Overdose Deaths") +
+
+# I want the cleanest theme.
+
 theme_classic()+
+
+# I changed the labelling on the x axis to be less cluttered.
+
 theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2020-04-02 ms_4.Rmd
moved drug overdose csv but graph still not showing

@@
 
-# This first data set is the new data set for the week. This data provides
+# This first data set provides
 # state-specific numbers from the CDC regarding the number of drug overdose

### 2020-04-02 ms_5.Rmd
moved drug overdose csv but graph still not showing

@@
 
-*Reading in drug overdose data was copy pasted from last week
-```{r, echo = FALSE}
+```{r}
 
-# This first data set is the new data set from last week. This data provides
+# This first data set provides
 # state-specific numbers from the CDC regarding the number of drug overdose
@@
 drugoverdosedeaths<- read.csv(
-  "VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
+  "ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
   clean_names() %>%
@@
  group_by("state_name")
+```
+
+```{r}
+# This data tracks the spending over a course of five years of a series of
+# drugs. I want to compare the spending levels of the different kinds of drugs
+# by categorizing them based on their use. Perhaps, we will find a correlation
+# between spending and drug type for opioid deaths.
+
+drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
+clean_names() %>%
+  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", 
+  "x2015_total_spending", "x2015_total_dosage_units", "x2016_total_spending", 
+  "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
+  "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
+```
+
+```{r}
+
+# This data set displays the number of deaths caused by opioids per county in NY
+# over a course of 15 years.
+
+deathbycounty_ny<- 
+  read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
+  clean_names()
+
+# Using data commons, I isolated drug spending in NY on some of the worst drugs
+# of addiction and treatment drugs such as Methadone, Buprenorphine, and Naloxone.
+
+drugspending_ny<- read.csv("New York Drug Spending.csv") %>%
+  clean_names() %>%
+  rename( Buprenorphine = "new_york_drug_dea_9064") %>%
+  rename( Cocaine = "new_york_drug_dea_9041l") %>%
+  rename( Methadone = "new_york_drug_dea_9250b") %>%
+  rename( Naloxone = "new_york_drug_dea_9411")
+
+# I want to compare the drug spending in NY (a state that has a lot of opioid
+# data with the country that has the worst opioid addiction rates, West
+# Virginia)
+
+drugspending_wv<- read.csv("West Virginia Drug Spending.csv") %>%
+clean_names() %>%
+  rename( Buprenorphine = "west_virginia_drug_dea_9064") %>%
+  rename( Cocaine = "west_virginia_drug_dea_9041l") %>%
+  rename( Methadone = "west_virginia_drug_dea_9250b") %>%
+  rename( Naloxone = "west_virginia_drug_dea_9411")
+
 ```
@@
 ```
+
+```{r}
+library(shiny)
+library(tidyverse)
+library(gt)
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+
+### Read in Data
+drugoverdosedeaths<- read.csv("ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv")
+
+drugoverdosedeaths <- drugoverdosedeaths %>%
+    clean_names() %>%
+    select("state", "year", "month", "indicator", 
+           "data_value", "state_name", "predicted_value") %>%
+    filter(str_detect(indicator, "Drug Overdose Deaths")) %>%
+    group_by("state_name") %>%
+    mutate(time = paste(month, year, sep = " "))
+drugoverdosedeaths$time<- as.factor(drugoverdosedeaths$time)
+drugoverdosedeaths$time<- fct_relevel(drugoverdosedeaths$time, "January 2015", "February 2015", 
+                                "March 2015", "April 2015", 
+                                "May 2015", "June 2015", "July 2015", "August 2015", "September 2015", 
+                                "October 2015", "November 2015", "December 2015", 
+                                "January 2016", "February 2016", "March 2016", 
+                                "April 2016", 
+                                "May 2016", "June 2016", "July 2016", "August 2016", "September 2016", 
+                                "October 2016", "November 2016", "December 2016", 
+                                "January 2017", "February 2017", 
+                                "March 2017", "April 2017", "May 2017", 
+                                "June 2017", "July 2017", "August 2017", "September 2017", 
+                                "October 2017", "November 2017", "December 2017", 
+                                "January 2018", "February 2018", 
+                                "March 2018", "April 2018", 
+                                "May 2018", "June 2018", "July 2018", "August 2018", 
+                                "September 2018", "October 2018", "November 2018", "December 2018",
+                                "January 2019", "February 2019", 
+                                "March 2019", "April 2019", 
+                                "May 2019", "June 2019", "July 2019") 
+drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
+
+        ggplot(drugoverdosedeaths, aes(x= reorder(state_name,-data_value), y= data_value)) +
+            geom_col() +
+            facet_wrap(~time)
+            labs(title = "Drug Overdose Deaths by State Over Time", 
+                 x = "State", 
+                 y = "Drug Overdose Deaths") +
+
+            theme_classic()+
+
+            theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```

### 2020-04-02 ms_6/app.R

@@ -1,0 +1,47 @@
+library(shiny)
+library(tidyverse)
+library(gt)
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+
+### Read in Data
+drugoverdosedeaths<- read.csv("VSRR_Provisional_Drug_Overdose_Death_Counts.csv")
+
+drugoverdosedeaths <- drugoverdosedeaths %>%
+    clean_names() %>%
+    select("state", "year", "month", "indicator", 
+           "data_value", "state_name", "predicted_value") %>%
+    filter(str_detect(indicator, "Drug Overdose Deaths")) %>%
+    group_by("state_name") %>%
+    mutate(time = paste(month, year, sep = " ")) %>%
+    mutate(time=mdy(time)) %>%
+    arrange(time)
+drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
+
+ui <- fluidPage(
+    titlePanel("Drug Overdoses by State Over Time"),
+    sliderInput("time", "Time Slider", min(drugoverdosedeaths$time),
+    max(drugoverdosedeaths$time),
+              value = min(drugoverdosedeaths$time)),
+    plotOutput("overdose_counts")
+)
+
+
+server <- function(input, output, session) {
+    #Rationing Map 
+    output$overdose_counts <- renderPlot({
+        ggplot(drugoverdosedeaths, aes(x= reorder(state_name,-data_value), y= data_value)) +
+            geom_col() +
+            facet_wrap(~time)
+            labs(title = "Drug Overdose Deaths by State Over Time", 
+                 x = "State", 
+                 y = "Drug Overdose Deaths") +
+            
+            theme_classic()+ 
+            
+            theme(axis.text.x = element_text(angle = 45, hjust = 1))
+    })
+    
+}
+shinyApp(ui, server)

### 2020-04-03 ms_6/app.R
shiny but i cant figure out how to change times on slider

@@ -4,6 +4,7 @@ library(gt)
 library(dplyr)
 library(ggplot2)
 library(lubridate)
+library(janitor)
 
 ### Read in Data
 drugoverdosedeaths<- read.csv("VSRR_Provisional_Drug_Overdose_Death_Counts.csv")
@@ -13,27 +14,30 @@ drugoverdosedeaths <- drugoverdosedeaths %>%
     select("state", "year", "month", "indicator", 
            "data_value", "state_name", "predicted_value") %>%
     filter(str_detect(indicator, "Drug Overdose Deaths")) %>%
-    group_by("state_name") %>%
+    filter(state_name != "United States") %>%
     mutate(time = paste(month, year, sep = " ")) %>%
     mutate(time=mdy(time)) %>%
+    group_by(time) %>%
     arrange(time)
 drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
+drugoverdosedeaths$predicted_value <-as.numeric(drugoverdosedeaths$predicted_value)
 
 ui <- fluidPage(
     titlePanel("Drug Overdoses by State Over Time"),
     sliderInput("time", "Time Slider", min(drugoverdosedeaths$time),
     max(drugoverdosedeaths$time),
-              value = min(drugoverdosedeaths$time)),
+              value = min(drugoverdosedeaths$time),
+              step = 1),
     plotOutput("overdose_counts")
 )
 
 
 server <- function(input, output, session) {
-    #Rationing Map 
     output$overdose_counts <- renderPlot({
-        ggplot(drugoverdosedeaths, aes(x= reorder(state_name,-data_value), y= data_value)) +
+       drugoverdosedeaths2<- drugoverdosedeaths %>%
+        filter(time == input$time)
+        ggplot(drugoverdosedeaths2, aes(x= reorder(state_name,-data_value), y= data_value)) +
             geom_col() +
-            facet_wrap(~time)
             labs(title = "Drug Overdose Deaths by State Over Time", 
                  x = "State", 
                  y = "Drug Overdose Deaths") +

### 2020-04-20 ms_5.Rmd
newdatasets and graphs!!

@@
 library(tidyverse)
+library(ggthemes)
+library(maps)
+library(leaflet)
 ```
@@
 
-### Read in Data
 drugoverdosedeaths<- read.csv("ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv")
@@
 
-        ggplot(drugoverdosedeaths, aes(x= reorder(state_name,-data_value), y= data_value)) +
+drugoverdosedeaths2 <- drugoverdosedeaths %>%
+ filter(time == "January 2015")
+
+        ggplot(drugoverdosedeaths2, aes(x= reorder(state_name,-data_value), y= data_value)) +
             geom_col() +
-            facet_wrap(~time)
             labs(title = "Drug Overdose Deaths by State Over Time", 
@@
 ```
+
+```{r, message=FALSE}
+#drug treatment centers
+
+treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
+  select(`Program Name`, Street, City, State, Zipcode)
+
+us_state <- map_data("state") %>%
+  rename(state_full = region)
+
+# https://simplemaps.com/data/us-cities
+
+longlatinfo <- read_csv("uscitieslonglat.csv") %>%
+  rename(City = city) %>%
+  rename(State = state_id)
+
+treatment_locations_map <- treatment_locations %>%
+  left_join(longlatinfo, by = c("City", "State")) %>%
+  group_by(State) %>%
+  mutate(count = n()) %>%
+  mutate(state_name = tolower(state_name)) %>%
+  rename(state_full = state_name) %>%
+  left_join(us_state, by = "state_full")
+
+ggplot(data = treatment_locations_map,
+            mapping = aes(x = long, y = lat.y,
+                          fill = count, group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,20,40,60,80,120)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom") +
+  labs(title = "Distribution of Opioid Treatment Centers in 2020",
+       color = "Number of Treatment Centers in State")
+
+# per capita pop data taken from https://www.census.gov/data/tables/time-series/demo/popest/2010s-state-total.html#par_textimage
+
+state_pop <- read_csv("State Population Estimate.csv") %>%
+  select(NAME, POPESTIMATE2019) %>%
+  rename(state_full = NAME) %>%
+  mutate(state_full = tolower(state_full))
+
+treatment_locations_map_per_cap <- treatment_locations_map %>%
+  left_join(state_pop, by = "state_full") %>%
+  group_by(State) %>%
+  mutate(count_per_pop = n()/POPESTIMATE2019)
+
+ggplot(data = treatment_locations_map_per_cap,
+            mapping = aes(x = long, y = lat.y,
+                          fill = count_per_pop, group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,20,40,60,80,120)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom") +
+  labs(title = "Distribution of Opioid Treatment Centers in 2020 (Per Capita)",
+       color = "Number of Treatment Centers in State Per Capita")
+```
+
+```{r, message = FALSE}
+#MA specific data
+
+madeathbyagegender <- read_csv("MAFatalOpioidOverdosesByAgeAndGender.csv")
+
+macountyheroin <- read_csv("MAHeroinAsPrimarySubstanceOfUse.csv")
+
+madeathbyrace <- read_csv("MAOpioidDeathsByRaceAndEthnicity.csv")
+
+matypeofsubstance <- read_csv("MAPrimarySubstanceofUseWhenEnteringTreatmentbyTown.csv")
+
+toxdeath <- read_csv("MAProportionDecedentsWithRxHistorybyToxicologyScreen.csv",
+         skip = 1)
+
+```
+
+```{r}
+mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv")
+
+mavsus_death <- mavsus_death %>%
+  group_by(Geography) %>%
+  filter (Year != "2015") %>%
+  filter (Year != "1999") %>%
+  rename(deathperhundredthousand =`Age-Adjusted Opioid-Related Death Rate per 100,000 People`)
+
+ggplot(mavsus_death, aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
+  geom_line() +
+theme_classic() +
+  labs(x = "Year",
+       y = "Deaths per 100,000",
+       title = "Comparison of Death Rates between Massachusetts and the United States",
+       subtitle = "Between the Years of 2000 and 2014") 
+```
+
+```{r}
+counties <- read_csv("zip_codes_states.csv") %>%
+  filter(state == "MA") %>%
+  rename(County = county) %>%
+  rename(Municipality = city)
+
+madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")
+
+countypop <- read_csv("countypop.csv") %>%
+  mutate(Pop = Pop/100000) %>%
+  rename(subregion = CTYNAME) %>%
+  mutate(subregion = tolower(subregion)) %>%
+  select(subregion, Pop)
+
+madeathbycountywlonglat <- madeathbycounty %>%
+  left_join(counties, by = "Municipality")
+
+madeathbycountywlonglat <- madeathbycountywlonglat %>%
+  select(County,
+         `Confirmed Opioid Related Death Count 2001-2005`,
+         `Confirmed Opioid Related Death Count 2006-2010`,
+         `Confirmed Opioid Related Death Count 2011-2015`,
+         `latitude`,
+         `longitude`,
+         `Municipality`) %>%
+  distinct(Municipality, .keep_all= TRUE) %>%
+  na.omit() %>%
+  rename(subregion = County) %>%
+  mutate(subregion = tolower(subregion)) %>%
+  group_by(subregion) %>%
+  mutate(total_deaths_2001.5 = sum(`Confirmed Opioid Related Death Count 2001-2005`)) %>%
+  mutate(total_deaths_2006.10 = sum(`Confirmed Opioid Related Death Count 2006-2010`)) %>%
+  mutate(total_deaths_2011.15 = sum(`Confirmed Opioid Related Death Count 2011-2015`)) %>%
+  distinct(subregion, .keep_all= TRUE)
+
+us_county <- map_data("county") %>%
+  filter(region == "massachusetts")
+
+# https://www.indexmundi.com/facts/united-states/quick-facts/massachusetts/percent-of-people-of-all-ages-in-poverty#table
+
+povertybycounty <- read_csv("PovertyByCounty.csv") %>%
+  rename(subregion = County) %>%
+  mutate(subregion = tolower(subregion))
+
+full_data <- us_county %>%
+  left_join(madeathbycountywlonglat, by = "subregion") %>%
+  left_join(countypop, by = "subregion") %>%
+  left_join(povertybycounty, by = "subregion") %>%
+  mutate(percap_2001.5 = total_deaths_2001.5/Pop) %>%
+  mutate(percap_2006.10 = total_deaths_2006.10/Pop) %>%
+  mutate(percap_2011.15 = total_deaths_2011.15/Pop)
+
+ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2001.5)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+
+full_data %>%
+  summarize(correlation = cor(percap_2001.5,`Poverty Percent`))
+
+ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2006.10)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+
+full_data %>%
+  summarize(correlation = cor(percap_2006.10,`Poverty Percent`))
+
+ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2011.15)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+
+full_data %>%
+  summarize(correlation = cor(percap_2011.15,`Poverty Percent`))
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = `Poverty Percent`,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,2,4,6,8,10)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = percap_2001.5,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,20,40,60,80,100)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = percap_2006.10,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,20,40,60,80,100)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = percap_2011.15,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,20,40,60,80,100)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+# don't put these in final app
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = total_deaths_2001.5,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,100,200,300,400,500,600)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = total_deaths_2006.10,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,100,200,300,400,500,600)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+ggplot(data = full_data,
+            mapping = aes(x = long, y = lat,
+                          fill = total_deaths_2011.15,
+                          group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A",
+                      breaks = c(0,100,200,300,400,500,600)) +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
+
+```
+
+```{r}
+
+#treatment centers in MA
+
+treatment_locations_map_mas <- treatment_locations %>%
+  left_join(longlatinfo, by = c("City", "State")) %>%
+  filter(State == "MA") %>%
+  drop_na()
+
+leaflet(options = leafletOptions(dragging = TRUE,
+                                 minZoom = 8, 
+                                 maxZoom = 10)) %>%
+  addProviderTiles("CartoDB") %>%
+  addCircleMarkers(data = treatment_locations_map_mas,
+                   radius = 3,
+                   label = ~`Program Name`)
+
+# need to figure this out bc color isnt working
+ # addLegend(position = "bottomright", colors = "#0000FF", title = "Opioid Treatment Sites")
+
+```

### 2020-04-24 ms_5.Rmd
gifs and new data sets/updated ms5 without extra ny/wv graphs

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
 
@@
 library(leaflet)
+library(gganimate)
 ```
@@
 
-# This first data set provides
-# state-specific numbers from the CDC regarding the number of drug overdose
-# deaths per state as well as the predicted numbers for drug overdose deaths. I
-# am super excited to plot this graph onto a map and look at how the worst
-# affected states change over time.
+# total deaths per year in u.s. by cancer, drug overdose, cvd
 
-drugoverdosedeaths<- read.csv(
-  "ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
-  clean_names() %>%
-  select("state", "year", "month", "indicator", 
-         "data_value", "state_name", "predicted_value") %>%
+deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", col_types = "ldddddd") %>%
+  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
+  mutate(`Proportion of 1999 Deaths by Cancer` = `Deaths by Cancer`/549829) %>%
+  mutate(`Proportion of 1999 Deaths by Drugs` = `Deaths by Drug Overdose`/19122) %>%
+  mutate(`Proportion of 1999 Deaths by CVD` = `Deaths by CVD`/949900)
 
-# I only want the information rows containing drug death information- I still
-# need to decipher the difference between "Deaths" and "Drug Overdose Deaths"
-# though.
+totaldeathsbycommondiseases <- deathsbycommondiseases %>%
+  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
+  pivot_longer(.,
+             cols = starts_with("Deaths by"),
+             names_prefix = "Deaths by",
+             values_to = "Deaths") %>%
+  rename(`Cause of Death` = name)
 
- filter(str_detect(indicator, "Death")) %>%
-
-# Ultimately, I want to look at state-specific data/trends so I want to group by
-# state.
-
- group_by("state_name")
+propdeathsbycommondiseases <- deathsbycommondiseases %>%
+  select(Year, `Proportion of 1999 Deaths by Cancer`, 
+         `Proportion of 1999 Deaths by Drugs`, 
+         `Proportion of 1999 Deaths by CVD`) %>%
+  pivot_longer(.,
+             cols = starts_with("Proportion of"),
+             names_prefix = "Proportion of 1999 Deaths by",
+             values_to = "Prop.Deaths") %>%
+  rename(`Cause of Death` = name)
 ```
@@
 ```{r}
-# This data tracks the spending over a course of five years of a series of
-# drugs. I want to compare the spending levels of the different kinds of drugs
-# by categorizing them based on their use. Perhaps, we will find a correlation
-# between spending and drug type for opioid deaths.
+plot_deathsbycommondiseases <- ggplot(totaldeathsbycommondiseases, aes(x = Year, y = Deaths, color = `Cause of Death`)) +
+  geom_line() +
+  scale_color_viridis_d() +
+  theme_classic() +
+   labs(x = "Year",
+       y = "Total Number of Deaths in the United States") +
+  geom_point() +
+  transition_reveal(Year)
 
-drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
-clean_names() %>%
-  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", 
-  "x2015_total_spending", "x2015_total_dosage_units", "x2016_total_spending", 
-  "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
-  "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
+anim_save("plot_deathsbycommondiseases.gif", plot_deathsbycommondiseases)
+
+prop_deathsbycommondiseases <- ggplot(propdeathsbycommondiseases, aes(x = Year, y = Prop.Deaths, color = `Cause of Death`)) +
+  geom_line() +
+  scale_color_viridis_d() +
+  theme_classic() +
+  labs(x = "Year",
+       y = "Proportion of 1999 Total Deaths in the United States") +
+  geom_point() +
+  transition_reveal(Year)
+
+anim_save("prop_deathsbycommondiseases.gif", prop_deathsbycommondiseases)
+
 ```
@@
 
-# This data set displays the number of deaths caused by opioids per county in NY
-# over a course of 15 years.
+# https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
-deathbycounty_ny<- 
-  read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
-  clean_names()
+death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv")
 
-# Using data commons, I isolated drug spending in NY on some of the worst drugs
-# of addiction and treatment drugs such as Methadone, Buprenorphine, and Naloxone.
+us_state <- map_data("state") %>%
+  rename(state_full = region)
 
-drugspending_ny<- read.csv("New York Drug Spending.csv") %>%
-  clean_names() %>%
-  rename( Buprenorphine = "new_york_drug_dea_9064") %>%
-  rename( Cocaine = "new_york_drug_dea_9041l") %>%
-  rename( Methadone = "new_york_drug_dea_9250b") %>%
-  rename( Naloxone = "new_york_drug_dea_9411")
+death_by_year_and_state <- death_by_year_and_state %>%
+  filter(`UCD - Drug/Alcohol Induced Code` == "D") %>%
+  rename(state_full = State) %>%
+  mutate(state_full = tolower(state_full)) %>%
+  left_join(us_state, by = "state_full" )
 
-# I want to compare the drug spending in NY (a state that has a lot of opioid
-# data with the country that has the worst opioid addiction rates, West
-# Virginia)
+death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
 
-drugspending_wv<- read.csv("West Virginia Drug Spending.csv") %>%
-clean_names() %>%
-  rename( Buprenorphine = "west_virginia_drug_dea_9064") %>%
-  rename( Cocaine = "west_virginia_drug_dea_9041l") %>%
-  rename( Methadone = "west_virginia_drug_dea_9250b") %>%
-  rename( Naloxone = "west_virginia_drug_dea_9411")
-
-```
-
-*The graphs below are new for this week. I don't know that I am satisfied with the readability of the last graph and will probably change it slightly in my final version.
-
-```{r, echo = FALSE}
-
-# For this week, I am going to plot the Drug Overdose Deaths in April of 2015 by
-# state according to the CDC. I want to use this data just because it is the
-# first recorded date of drug overdose deaths. I am hoping there will be some
-# expected trend across states.
-
-# I am not sure if this makes a difference but I am going to change the variable
-# to be numeric (from factor) to rank later on.
-
-drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
-
-# Here, I want to rank the deaths by state in descending order and filter them
-# such that we are only looking at info from April of 2015.
-
-drugodapr15<- drugoverdosedeaths %>%
-  arrange(desc(data_value)) %>%
-  select(state_name, `"state_name"`, data_value, indicator, month, year) %>%
-  filter(year == "2015",
-        month == "April",
-        indicator == "Number of Drug Overdose Deaths")
-
-# Now, I want to ggplot the data. In the aesthetics, when I simply plot the
-# state name vs data value, the data values do not rank in descending order
-# despite specifying earlier, so I used the trick from PSet 3 of reordering in
-# the x specifications.
-
-ggplot(drugodapr15, aes(x= reorder(state_name,-data_value), y= data_value)) +
-geom_col() +
-
-# I want to label the title and x and y axes so that they are more descriptive.
-
-labs(title = "Drug Overdose Deaths by State (April 2015)", 
-     x = "State", 
-     y = "Drug Overdose Deaths") +
-
-# I changed the theme such that the backing of the graph is blank.
-
-theme_classic()+
-
-# The x axis details are cluttered so I chose to angle them for easy reading.
-
-theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-```{r}
-
-# I am going to do the same with July 2019, which is the latest date of
-# recording. All code is the same except that the filter specifications at the
-# start are for July of 2019 and I changed the title to specify July of 2019.
-
-drugodjul19<- drugoverdosedeaths %>%
-  arrange(desc(data_value)) %>%
-  select(state_name, `"state_name"`, data_value, indicator, month, year) %>%
-  filter(year == "2019",
-        month == "July",
-        indicator == "Number of Drug Overdose Deaths")
-
-ggplot(drugodjul19, aes(x= reorder(state_name,-data_value), y= data_value)) +
-geom_col() +
-labs(title = "Drug Overdose Deaths by State (July 2019)", 
-     x = "State", 
-     y = "Drug Overdose Deaths") +
-theme_classic()+ 
-theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-```{r, echo = FALSE}
-
-# I now want to look at how drug overdose death levels changed between 2019 and
-# 2015 by state. To do so, I combined the two data tables from above using a
-# full join.
-
-difference_over_time<- drugodapr15 %>%
-  full_join(drugodjul19, by = "state_name") %>%
-
-# I also want to add a column which indicates the difference in deaths between
-# the two data tables.
-
-mutate(difference = data_value.y - data_value.x)
-
-# I want to graph the difference over time by state with the x axis ranking the
-# difference decrease.
-
-ggplot(difference_over_time, aes(x= reorder(state_name,-difference), 
-                                 y= difference)) +
-geom_col() +
-
-# I also changed the title label and the y axis label to specify the
-# differences.
-
-labs(title = 
-    "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
-     x = "State", 
-     y = "Difference in Drug Overdose Deaths") +
-
-# I want the cleanest theme.
-
-theme_classic()+
-
-# I changed the labelling on the x axis to be less cluttered.
-
-theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-```{r}
-library(shiny)
-library(tidyverse)
-library(gt)
-library(dplyr)
-library(ggplot2)
-library(lubridate)
-
-drugoverdosedeaths<- read.csv("ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv")
-
-drugoverdosedeaths <- drugoverdosedeaths %>%
-    clean_names() %>%
-    select("state", "year", "month", "indicator", 
-           "data_value", "state_name", "predicted_value") %>%
-    filter(str_detect(indicator, "Drug Overdose Deaths")) %>%
-    group_by("state_name") %>%
-    mutate(time = paste(month, year, sep = " "))
-drugoverdosedeaths$time<- as.factor(drugoverdosedeaths$time)
-drugoverdosedeaths$time<- fct_relevel(drugoverdosedeaths$time, "January 2015", "February 2015", 
-                                "March 2015", "April 2015", 
-                                "May 2015", "June 2015", "July 2015", "August 2015", "September 2015", 
-                                "October 2015", "November 2015", "December 2015", 
-                                "January 2016", "February 2016", "March 2016", 
-                                "April 2016", 
-                                "May 2016", "June 2016", "July 2016", "August 2016", "September 2016", 
-                                "October 2016", "November 2016", "December 2016", 
-                                "January 2017", "February 2017", 
-                                "March 2017", "April 2017", "May 2017", 
-                                "June 2017", "July 2017", "August 2017", "September 2017", 
-                                "October 2017", "November 2017", "December 2017", 
-                                "January 2018", "February 2018", 
-                                "March 2018", "April 2018", 
-                                "May 2018", "June 2018", "July 2018", "August 2018", 
-                                "September 2018", "October 2018", "November 2018", "December 2018",
-                                "January 2019", "February 2019", 
-                                "March 2019", "April 2019", 
-                                "May 2019", "June 2019", "July 2019") 
-drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
-
-drugoverdosedeaths2 <- drugoverdosedeaths %>%
- filter(time == "January 2015")
-
-        ggplot(drugoverdosedeaths2, aes(x= reorder(state_name,-data_value), y= data_value)) +
-            geom_col() +
-            labs(title = "Drug Overdose Deaths by State Over Time", 
-                 x = "State", 
-                 y = "Drug Overdose Deaths") +
-
-            theme_classic()+
-
-            theme(axis.text.x = element_text(angle = 45, hjust = 1))
+ggplot(data = death_by_year_and_state,
+            mapping = aes(x = long, y = lat,
+                          fill = `Crude Rate`, group = group)) + 
+  geom_polygon(color = "gray90", size = 0.05) + 
+  theme_map() +
+  scale_fill_gradient(low = "white", high = "#CB454A") +
+    guides(fill = guide_legend(nrow = 1)) + 
+    theme(legend.position = "bottom")
 
@@
 
-us_state <- map_data("state") %>%
-  rename(state_full = region)
+# state populations
+
+state_pop <- read_csv("State Population Estimate.csv") %>%
+  rename(state_full = NAME) %>%
+  mutate(state_full = tolower(state_full))
 
@@
        color = "Number of Treatment Centers in State")
-
-# per capita pop data taken from https://www.census.gov/data/tables/time-series/demo/popest/2010s-state-total.html#par_textimage
-
-state_pop <- read_csv("State Population Estimate.csv") %>%
-  select(NAME, POPESTIMATE2019) %>%
-  rename(state_full = NAME) %>%
-  mutate(state_full = tolower(state_full))
 
@@
 ```{r}
-mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv")
+mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", col_types = "cdc")
 
@@
 
-ggplot(mavsus_death, aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
+plot_mavsus_death <- ggplot(mavsus_death, aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
   geom_line() +
+  scale_color_viridis_d() +
 theme_classic() +
@@
        title = "Comparison of Death Rates between Massachusetts and the United States",
-       subtitle = "Between the Years of 2000 and 2014") 
+       subtitle = "Between the Years of 2000 and 2014") +
+  geom_point() +
+  transition_reveal(Year)
+
+anim_save("plot_mavsus_death.gif", plot_mavsus_death)
+
 ```

### 2020-04-24 app.R

@@ -1,0 +1,165 @@
+library(shiny)
+library(tidyverse)
+library(gt)
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+library(janitor)
+library(shinyWidgets)
+library(shinythemes)
+library(ggthemes)
+library(maps)
+
+### Read in Data
+
+# https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
+
+death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv")
+
+us_state <- map_data("state") %>%
+  rename(state_full = region)
+
+death_by_year_and_state <- death_by_year_and_state %>%
+  filter(`UCD - Drug/Alcohol Induced Code` == "D") %>%
+  rename(state_full = State) %>%
+  mutate(state_full = tolower(state_full)) %>%
+  left_join(us_state, by = "state_full" ) %>%
+  drop_na(`Crude Rate`)
+
+death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
+
+ui <- navbarPage(theme = shinytheme("flatly"),
+                 "Opioid Trends Across the United States",
+                 tabPanel("About",
+                          column(7,
+                                 h1("Background"),
+                                 p("This project is meant to look into trends 
+                                  regarding the issue of opioid overuse. In the
+                                   1990's doctors overprescribed pain relievers 
+                                   as a temporary bandage to disease without 
+                                   knowing the potential for addiction. 
+                                   Since then, the misuse of opioids has ignited 
+                                   a national crisis that has contributed to 
+                                   worsening health outcomes for thousands and 
+                                   their families throughout the United States. 
+                                   Opioid misuse is not only an emotional burden 
+                                   for patients and their families, 
+                                   it also loses the United States billions of 
+                                   dollars yearly. Thus, this study looks into 
+                                   some of the spending trends of opioid use in 
+                                   conjunction with death rates."),
+                                 
+                                 p("The plan for this project is to examine the 
+                                   distribution of opioid deaths across states
+                                   over the past twenty years. Having looked at 
+                                   this data, I also sought to look at the 
+                                   distribution of treatment centers to see if 
+                                   the treatment center distribution matches up 
+                                   with opioid death distribution: are the 
+                                   states that need the most attention getting 
+                                   the most attention?"),
+                                 
+                                 p("Following this, I analyzed the data for 
[…]

### 2020-04-25 ms_5.Rmd
models and new data

@@
 library(gganimate)
+library(broom)
 ```
@@
 
+# not in final
+
 ggplot(data = treatment_locations_map,
@@
 
+```
+
+```{r}
+# in final
+
 ggplot(data = treatment_locations_map_per_cap,
@@
        color = "Number of Treatment Centers in State Per Capita")
+
 ```
@@
 ```
+
+```{r}
+
+# models generated from data from kff.org
+
+# ageopioid
+
+ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", col_types = "dcc") %>%
+  clean_names() %>%
+  na.omit() %>%
+  filter(age_range != "Total") %>%
+  mutate(age_range = as.factor(age_range)) %>%
+    mutate(number_of_deaths = as.numeric(number_of_deaths))
+
+ggplot(ageopioidmodel, aes(year, number_of_deaths, color = age_range)) +
+  geom_point() +
+  scale_color_viridis_d() +
+  theme_classic() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Age Range")
+
+#raceopioid
+
+raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
+  clean_names() %>%
+  na.omit() %>%
+  mutate(race = as.factor(race)) %>%
+  mutate(opioid_deaths = as.numeric(opioid_deaths))
+
+ggplot(raceopioidmodel, aes(year, opioid_deaths, color = race)) +
+  geom_point() +
+  scale_color_viridis_d() +
+  theme_classic() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Race")
+```
+
+```{r}
+
+# filtering data by race/age
+ages0_24 <- ageopioidmodel %>%
+  filter(age_range == "0-24")
+
+ages0_24 %>%
+    lm(number_of_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+
+ages25_34 <- ageopioidmodel %>%
+  filter(age_range == "25-34")
+
+ages25_34 %>%
+    lm(number_of_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+
+ages35_44 <- ageopioidmodel %>%
+  filter(age_range == "35-44")
+
+ages35_44 %>%
+    lm(number_of_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+
+ages45_54 <- ageopioidmodel %>%
+  filter(age_range == "45-54")
+
+ages45_54 %>%
+    lm(number_of_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+
+ages55_up <- ageopioidmodel %>%
+  filter(age_range == "55+")
+
+ages55_up %>%
+    lm(number_of_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+
+# averageageopioidmodel <- ageopioidmodel %>%
+  # drop_na() %>%
+  # group_by(age_range) %>%
+  # summarize(mean_deaths = mean(number_of_deaths))
+
+# averageageopioid_model <- lm(mean_deaths ~ age_range, data = averageageopioidmodel)
+# averageageopioid_model %>%
+  # tidy(conf.int = TRUE) %>%
+  # select(term, estimate, conf.low, conf.high)
+```

### 2020-04-25 app.R

@@ -9,6 +9,7 @@ library(shinyWidgets)
 library(shinythemes)
 library(ggthemes)
 library(maps)
+library(leaflet)
 
 ### Read in Data
 
@@ -28,26 +29,100 @@ death_by_year_and_state <- death_by_year_and_state %>%
 
 death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
 
+treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
+  select(`Program Name`, Street, City, State, Zipcode)
+
+# state populations
+
+state_pop <- read_csv("State Population Estimate.csv") %>%
+  rename(state_full = NAME) %>%
+  mutate(state_full = tolower(state_full))
+
+# https://simplemaps.com/data/us-cities
+
+longlatinfo <- read_csv("uscitieslonglat.csv") %>%
+  rename(City = city) %>%
+  rename(State = state_id)
+
+treatment_locations_map <- treatment_locations %>%
+  left_join(longlatinfo, by = c("City", "State")) %>%
+  group_by(State) %>%
+  mutate(count = n()) %>%
+  mutate(state_name = tolower(state_name)) %>%
+  rename(state_full = state_name) %>%
+  left_join(us_state, by = "state_full") 
+
+treatment_locations_map_per_cap <- treatment_locations_map %>%
+  left_join(state_pop, by = "state_full") %>%
+  group_by(State) %>%
+  mutate(count_per_pop = n()/POPESTIMATE2019)
+
+treatment_locations_map_mas <- treatment_locations %>%
+  left_join(longlatinfo, by = c("City", "State")) %>%
+  filter(State == "MA") %>%
+  drop_na()
+
+counties <- read_csv("zip_codes_states.csv") %>%
+  filter(state == "MA") %>%
+  rename(County = county) %>%
+  rename(Municipality = city)
+
+madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")
+
+countypop <- read_csv("countypop.csv") %>%
+  mutate(Pop = Pop/100000) %>%
+  rename(subregion = CTYNAME) %>%
+  mutate(subregion = tolower(subregion)) %>%
+  select(subregion, Pop)
+
+madeathbycountywlonglat <- madeathbycounty %>%
+  left_join(counties, by = "Municipality")
+
+madeathbycountywlonglat <- madeathbycountywlonglat %>%
[…]

### 2020-04-25 ms_6/app.R

@@ -1,53 +1,0 @@
-library(shiny)
-library(tidyverse)
-library(gt)
-library(dplyr)
-library(ggplot2)
-library(lubridate)
-library(janitor)
-library(shinyWidgets)
-
-### Read in Data
-drugoverdosedeaths<- read.csv("VSRR_Provisional_Drug_Overdose_Death_Counts.csv")
-
-drugoverdosedeaths <- drugoverdosedeaths %>%
-  clean_names() %>%
-  select("state", "year", "month", "indicator", 
-         "data_value", "state_name", "predicted_value") %>%
-  filter(str_detect(indicator, "Drug Overdose Deaths")) %>%
-  filter(state_name != "United States") %>%
-  mutate(time = paste(month, year, sep = " ")) %>%
-  mutate(time=mdy(time) - 19) %>%
-  mutate(month_number = month(time)) %>%
-    group_by(time) %>%
-    arrange(time)
-drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)
-drugoverdosedeaths$predicted_value <-as.numeric(drugoverdosedeaths$predicted_value)
-
-ui <- fluidPage(
-  titlePanel("Drug Overdoses by State Over Time"),
-  sliderTextInput("time", "Time", 
-                            from_min = as.Date("2015-01-01"),
-                            from_max =as.Date("2019-07-01"),
-                            timeFormat="%b %Y",
-                            choices = levels(as.factor(drugoverdosedeaths$time))),
-                plotOutput("overdose_counts")
-)
-
-server <- function(input, output, session) {
-    output$overdose_counts <- renderPlot({
-       drugoverdosedeaths2<- drugoverdosedeaths %>%
-        filter(time == input$time)
-        ggplot(drugoverdosedeaths2, aes(x= state_name, y= data_value)) +
-            geom_col() +
-            labs(title = "Drug Overdose Deaths by State Over Time", 
-                 x = "State", 
-                 y = "Drug Overdose Deaths") +
-            
-            theme_classic()+ 
-            
-            theme(axis.text.x = element_text(angle = 45, hjust = 1))
-    })
-    
-}
-shinyApp(ui, server)

### 2020-04-26 ms_5.Rmd
just need to fix legends in app/titles

@@
 
-ggplot(ageopioidmodel, aes(year, number_of_deaths, color = age_range)) +
+raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
+  clean_names() %>%
+  na.omit() %>%
+  mutate(race = as.factor(race)) %>%
+  mutate(opioid_deaths = as.numeric(opioid_deaths))
+
+ageandracemodel <- ageopioidmodel %>%
+  na.omit() %>%
+  full_join(raceopioidmodel, by = "year") %>%
+  rename(deathsbyage = number_of_deaths) %>%
+  rename(deathsbyrace = opioid_deaths)
+
+ggplot(ageandracemodel, aes(year, deathsbyage, color = age_range)) +
   geom_point() +
@@
 
-#raceopioid
-
-raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
-  clean_names() %>%
-  na.omit() %>%
-  mutate(race = as.factor(race)) %>%
-  mutate(opioid_deaths = as.numeric(opioid_deaths))
-
-ggplot(raceopioidmodel, aes(year, opioid_deaths, color = race)) +
+ggplot(ageandracemodel, aes(year, deathsbyrace, color = race)) +
   geom_point() +
@@
 ```{r}
+# filtering data by race to show effects of time on deaths in each race group
 
-# filtering data by race/age
+death_vs_white <- raceopioidmodel %>%
+  filter(race == "White, Non-Hispanic")
+
+death_vs_white %>%
+    lm(opioid_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high) %>%
+    rename(Coefficient = estimate) %>%
+    rename(`Lower End` = conf.low) %>%
+    rename(`Upper End` = conf.high) %>%
+    gt()
+
+death_vs_black <- raceopioidmodel %>%
+  filter(race == "Black, Non-Hispanic")
+
+death_vs_black %>%
+    lm(opioid_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+
+death_vs_hispanic <- raceopioidmodel %>%
+  filter(race == "Hispanic")
+
+death_vs_hispanic %>%
+    lm(opioid_deaths ~ year, data = .) %>% 
+    tidy(conf.int = TRUE) %>% 
+    select(term, estimate, conf.low, conf.high)
+```
+
+```{r}
+
+# filtering data by age to show effects of time on deaths in each age group
+
 ages0_24 <- ageopioidmodel %>%
@@
 
-# averageageopioidmodel <- ageopioidmodel %>%
-  # drop_na() %>%
-  # group_by(age_range) %>%
-  # summarize(mean_deaths = mean(number_of_deaths))
+# looking at average deaths in each age group
 
-# averageageopioid_model <- lm(mean_deaths ~ age_range, data = averageageopioidmodel)
-# averageageopioid_model %>%
-  # tidy(conf.int = TRUE) %>%
-  # select(term, estimate, conf.low, conf.high)
+averageageopioidbyage <- ageopioidmodel %>%
+  drop_na() %>%
+  group_by(age_range) %>%
+  summarize(mean_deaths = mean(number_of_deaths))
+
+# showing effects of being in each age group on number of deaths idt this makes sense
+
+averageageopioid_model <- lm(number_of_deaths ~ age_range, data = ageopioidmodel)
+averageageopioid_model %>%
+  tidy(conf.int = TRUE) %>%
+  select(term, estimate, conf.low, conf.high)
 ```

### 2020-04-26 app.R

@@ -10,6 +10,9 @@ library(shinythemes)
 library(ggthemes)
 library(maps)
 library(leaflet)
+library(plotly)
+library(DT)
+library(broom)
 
 ### Read in Data
 
@@ -113,16 +116,66 @@ full_data <- us_county %>%
   mutate(percap_2006.10 = total_deaths_2006.10/Pop) %>%
   mutate(percap_2011.15 = total_deaths_2011.15/Pop) 
 
+ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", col_types = "dcc") %>%
+  clean_names() %>%
+  na.omit() %>%
+  filter(age_range != "Total") %>%
+  mutate(age_range = as.factor(age_range)) %>%
+  mutate(number_of_deaths = as.numeric(number_of_deaths))
+
+raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
+  clean_names() %>%
+  na.omit() %>%
+  mutate(race = as.factor(race)) %>%
+  mutate(opioid_deaths = as.numeric(opioid_deaths))
+
+ageandracemodel <- ageopioidmodel %>%
+  na.omit() %>%
+  full_join(raceopioidmodel, by = "year") %>%
+  rename(deathsbyage = number_of_deaths) %>%
+  rename(deathsbyrace = opioid_deaths)
+
+deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", col_types = "ldddddd") %>%
+  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
+  mutate(`Proportion of 1999 Deaths by Cancer` = `Deaths by Cancer`/549829) %>%
+  mutate(`Proportion of 1999 Deaths by Drugs` = `Deaths by Drug Overdose`/19122) %>%
+  mutate(`Proportion of 1999 Deaths by CVD` = `Deaths by CVD`/949900)
+
+totaldeathsbycommondiseases <- deathsbycommondiseases %>%
+  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
+  pivot_longer(.,
+               cols = starts_with("Deaths by"),
+               names_prefix = "Deaths by",
+               values_to = "Deaths") %>%
+  rename(`Cause of Death` = name)
+
+propdeathsbycommondiseases <- deathsbycommondiseases %>%
[…]

### 2020-04-26 ms_5.Rmd
still getting col spec errors

@@
 
-anim_save("plot_deathsbycommondiseases.gif", plot_deathsbycommondiseases)
-
 prop_deathsbycommondiseases <- ggplot(propdeathsbycommondiseases, aes(x = Year, y = Prop.Deaths, color = `Cause of Death`)) +
@@
 
-anim_save("prop_deathsbycommondiseases.gif", prop_deathsbycommondiseases)
-
 ```
@@
 
-death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv")
+death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
+                                    col_types = "lcdccddddddddic")
 
@@
   transition_reveal(Year)
-
-anim_save("plot_mavsus_death.gif", plot_mavsus_death)
 

### 2020-04-26 app.R

@@ -18,7 +18,8 @@ library(broom)
 
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
-death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv")
+death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
+                                    col_types = "lcdccddddd")
 
 us_state <- map_data("state") %>%
   rename(state_full = region)
@@ -37,13 +38,13 @@ treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
 
 # state populations
 
-state_pop <- read_csv("State Population Estimate.csv") %>%
+state_pop <- read_csv("State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
   rename(state_full = NAME) %>%
   mutate(state_full = tolower(state_full))
 
 # https://simplemaps.com/data/us-cities
 
-longlatinfo <- read_csv("uscitieslonglat.csv") %>%
+longlatinfo <- read_csv("uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
   rename(City = city) %>%
   rename(State = state_id)
 
@@ -65,14 +66,14 @@ treatment_locations_map_mas <- treatment_locations %>%
   filter(State == "MA") %>%
   drop_na()
 
-counties <- read_csv("zip_codes_states.csv") %>%
+counties <- read_csv("zip_codes_states.csv", col_types = "dddccc") %>%
   filter(state == "MA") %>%
   rename(County = county) %>%
   rename(Municipality = city)
 
 madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")
 
-countypop <- read_csv("countypop.csv") %>%
+countypop <- read_csv("countypop.csv", col_types = "cdcdddc") %>%
   mutate(Pop = Pop/100000) %>%
   rename(subregion = CTYNAME) %>%
   mutate(subregion = tolower(subregion)) %>%
@@ -104,7 +105,7 @@ us_county <- map_data("county") %>%
 
[…]

### 2020-04-26 app.R
forgot to push

@@ -14,7 +14,7 @@ library(plotly)
 library(DT)
 library(broom)
 
-### Read in Data
+# Read in Data
 
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 

### 2020-04-26 app.R
fixing col spec

@@ -19,7 +19,7 @@ library(broom)
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
 death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
-                                    col_types = "lcdccddddd")
+                                    col_types = "lcdccddddc")
 
 us_state <- map_data("state") %>%
   rename(state_full = region)

### 2020-04-26 ms_5.Rmd
fixing col spec
fixed col spec except for death by year and state

@@
 death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
-                                    col_types = "lcdccddddddddic")
+                                    col_types = "lcdccddddc")
 
@@
 
-treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
+treatment_locations <- read_csv("Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
   select(`Program Name`, Street, City, State, Zipcode)
@@
 
-state_pop <- read_csv("State Population Estimate.csv") %>%
+state_pop <- read_csv("State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
   rename(state_full = NAME) %>%
@@
 
-longlatinfo <- read_csv("uscitieslonglat.csv") %>%
+longlatinfo <- read_csv("uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
   rename(City = city) %>%
@@
 
-madeathbyagegender <- read_csv("MAFatalOpioidOverdosesByAgeAndGender.csv")
+madeathbyagegender <- read_csv("MAFatalOpioidOverdosesByAgeAndGender.csv",
+                               col_types = "ccddd")
 
-macountyheroin <- read_csv("MAHeroinAsPrimarySubstanceOfUse.csv")
+macountyheroin <- read_csv("MAHeroinAsPrimarySubstanceOfUse.csv",
+                           col_types = "ccccc")
 
-madeathbyrace <- read_csv("MAOpioidDeathsByRaceAndEthnicity.csv")
+madeathbyrace <- read_csv("MAOpioidDeathsByRaceAndEthnicity.csv",
+                          col_types = "dcdd")
 
-matypeofsubstance <- read_csv("MAPrimarySubstanceofUseWhenEnteringTreatmentbyTown.csv")
+matypeofsubstance <- read_csv("MAPrimarySubstanceofUseWhenEnteringTreatmentbyTown.csv",
+                              col_types = "cccccccccccc")
 
 toxdeath <- read_csv("MAProportionDecedentsWithRxHistorybyToxicologyScreen.csv",
-         skip = 1)
+         skip = 1,
+         col_types = "cdccccccdc")
 
@@
 ```{r}
-counties <- read_csv("zip_codes_states.csv") %>%
+counties <- read_csv("zip_codes_states.csv",
+                     col_types = "dddccc") %>%
   filter(state == "MA") %>%
@@
 
-madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")
+madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv",
+                            col_types = "cdcdddc")
 
-countypop <- read_csv("countypop.csv") %>%
+countypop <- read_csv("countypop.csv",
+                      col_types = "cd") %>%
   mutate(Pop = Pop/100000) %>%
@@
 
-povertybycounty <- read_csv("PovertyByCounty.csv") %>%
+povertybycounty <- read_csv("PovertyByCounty.csv",
+                            col_types = "cd") %>%
   rename(subregion = County) %>%

### 2020-04-26 app.R

@@ -71,7 +71,8 @@ counties <- read_csv("zip_codes_states.csv", col_types = "dddccc") %>%
   rename(County = county) %>%
   rename(Municipality = city)
 
-madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")
+madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv", 
+                            col_types = "cdcdddc")
 
 countypop <- read_csv("countypop.csv", col_types = "cdcdddc") %>%
   mutate(Pop = Pop/100000) %>%

### 2020-04-26 app.R
cleaned data for drug overdose deaths/ got rid of gifs

@@ -33,7 +33,7 @@ death_by_year_and_state <- death_by_year_and_state %>%
 
 death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
 
-treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
+treatment_locations <- read_csv("Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
   select(`Program Name`, Street, City, State, Zipcode)
 
 # state populations

### 2020-04-26 ms_5.Rmd
fixed col spec- only error is NAs introcued by coercion

@@
 death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
-                                    col_types = "lcdccddddc")
+                                    col_types = "lcdccddddc") %>%
+  filter(`Crude Rate` != "Unreliable")
 
@@
 countypop <- read_csv("countypop.csv",
-                      col_types = "cd") %>%
+                      col_types = "cdd") %>%
   mutate(Pop = Pop/100000) %>%

### 2020-04-26 app.R

@@ -19,7 +19,8 @@ library(broom)
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
 death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
-                                    col_types = "lcdccddddc")
+                                    col_types = "lcdccddddc") %>%
+  filter(`Crude Rate` != "Unreliable")
 
 us_state <- map_data("state") %>%
   rename(state_full = region)
@@ -67,14 +68,17 @@ treatment_locations_map_mas <- treatment_locations %>%
   drop_na()
 
 counties <- read_csv("zip_codes_states.csv", col_types = "dddccc") %>%
+  na.omit() %>%
   filter(state == "MA") %>%
   rename(County = county) %>%
   rename(Municipality = city)
 
 madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv", 
-                            col_types = "cdcdddc")
+                            col_types = "cdcdddc") %>%
+  na.omit()
 
-countypop <- read_csv("countypop.csv", col_types = "cdcdddc") %>%
+countypop <- read_csv("countypop.csv", col_types = "cdd") %>%
+  na.omit() %>%
   mutate(Pop = Pop/100000) %>%
   rename(subregion = CTYNAME) %>%
   mutate(subregion = tolower(subregion)) %>%
@@ -107,6 +111,7 @@ us_county <- map_data("county") %>%
 # https://www.indexmundi.com/facts/united-states/quick-facts/massachusetts/percent-of-people-of-all-ages-in-poverty#table
 
 povertybycounty <- read_csv("PovertyByCounty.csv", col_types = "cd") %>%
+  na.omit() %>%
   rename(subregion = County) %>%
   mutate(subregion = tolower(subregion))
 
@@ -138,6 +143,7 @@ ageandracemodel <- ageopioidmodel %>%
   rename(deathsbyrace = opioid_deaths)
 
 deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", col_types = "ldddddd") %>%
+  na.omit() %>%
   select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
[…]

### 2020-04-26 app.R
unnecessary data removed

@@ -125,7 +125,7 @@ full_data <- us_county %>%
 
 ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", col_types = "dcc") %>%
   clean_names() %>%
-  na.omit() %>%
+  filter(number_of_deaths != "N/A") %>%
   filter(age_range != "Total") %>%
   mutate(age_range = as.factor(age_range)) %>%
   mutate(number_of_deaths = as.numeric(number_of_deaths))
@@ -143,7 +143,6 @@ ageandracemodel <- ageopioidmodel %>%
   rename(deathsbyrace = opioid_deaths)
 
 deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", col_types = "ldddddd") %>%
-  na.omit() %>%
   select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
   mutate(`Proportion of 1999 Deaths by Cancer` = `Deaths by Cancer`/549829) %>%
   mutate(`Proportion of 1999 Deaths by Drugs` = `Deaths by Drug Overdose`/19122) %>%

### 2020-04-26 app.R
switched tab order

@@ -174,56 +174,6 @@ mavsus_death <- mavsus_death %>%
 
 ui <- navbarPage(theme = shinytheme("flatly"),
                  "Opioid Trends Across the United States",
-                 tabPanel("About",
-                          column(9,
-                                 h1("Background"),
-                                 p("Opioids have been prescribed for pain management for many decades, 
-                                 although they are effective at treating acute, not chronic pain. 
-                                   In treating chronic pain, opioids can be more harmful than useful. 
-                                   Opioids are cheap, but have high rates of addiction and can lead to 
-                                   dependence and ultimately death from overdose.Since the 1990s, the United States
-                                   has experienced a crisis created by overprescription of opioids. Thousands die yearly
-                                   from uncontrollable overuse."),
-                                 
-                                 p("The plan for this project is to examine the 
-                                   distribution of opioid deaths across states
-                                   over the past twenty years. Having looked at 
-                                   this data, I also sought to look at the 
-                                   distribution of treatment centers to see if 
-                                   the treatment center distribution matches up 
-                                   with opioid death distribution: are the 
-                                   states that need the most attention getting 
-                                   the most attention?"),
-                                 
-                                 p("Following this, I analyzed the data for 
-                                   the state nearest and dearest to our hearts, 
-                                   Massachusetts. How does the distribution 
-                                   of opioid deaths in Massachusetts look and 
-                                   how are treatment centers distributed 
-                                   according to deaths per capita in 
-                                   different counties?"),
-                                 
-                                 p("Finally, I sought to create a model between factors of interest such as 
-                                   age and ethnicity and increases in opioid deaths over the past twenty years."),
-                                 h1("The Data"),
-                                 p("Data concerning the entire United States came from the CDC's Wonder Archive. Massachusetts-specific data came from
-                                   chapter55.digital.mass.gov."),
-                                 p("My project code can be found on my",
-                                   a("GitHub",
-                                     href = "https://github.com/suruchiramanujan/final-project.git",)),
-                                 h1("About Me"),
-                                 p("My name is Suruchi Ramanujan and I am a 
-                                   senior in Quincy House studying 
-                                   Molecular and Cellular Biology. In my free time, 
-                                   I enjoy performing Indian Classical dance, running, 
-                                   and grinding through my DataCamp exercises."),
[…]

### 2020-04-26 all_data.R
all data converted to rds to reduce space

@@ -1,0 +1,200 @@
+library(shiny)
+library(tidyverse)
+library(gt)
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+library(janitor)
+library(shinyWidgets)
+library(shinythemes)
+library(ggthemes)
+library(maps)
+library(leaflet)
+library(plotly)
+library(DT)
+library(broom)
+
+# https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
+
+death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
+                                    col_types = "lcdccddddc") %>%
+  filter(`Crude Rate` != "Unreliable")
+
+us_state <- map_data("state") %>%
+  rename(state_full = region)
+
+death_by_year_and_state <- death_by_year_and_state %>%
+  filter(`UCD - Drug/Alcohol Induced Code` == "D") %>%
+  rename(state_full = State) %>%
+  mutate(state_full = tolower(state_full)) %>%
+  left_join(us_state, by = "state_full" ) %>%
+  drop_na(`Crude Rate`)
+
+death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
+
+write_rds(death_by_year_and_state, "death_by_year_and_state.rds")
+
+treatment_locations <- read_csv("Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
+  select(`Program Name`, Street, City, State, Zipcode)
+
+write_rds(treatment_locations, "treatment_locations.rds")
+
+# state populations
+
+state_pop <- read_csv("State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
+  rename(state_full = NAME) %>%
+  mutate(state_full = tolower(state_full))
+
+write_rds(state_pop, "state_pop.rds")
+
+# https://simplemaps.com/data/us-cities
+
+longlatinfo <- read_csv("uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
+  rename(City = city) %>%
+  rename(State = state_id)
+
+write_rds(longlatinfo, "longlatinfo.rds")
+
[…]

### 2020-04-26 app.R
all code working- still need to fix legends

@@ -18,71 +18,44 @@ library(broom)
 
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
-death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
-                                    col_types = "lcdccddddc") %>%
-  filter(`Crude Rate` != "Unreliable")
+death_by_year_and_state <- read_rds("death_by_year_and_state.rds")
 
-us_state <- map_data("state") %>%
-  rename(state_full = region)
-
-death_by_year_and_state <- death_by_year_and_state %>%
-  filter(`UCD - Drug/Alcohol Induced Code` == "D") %>%
-  rename(state_full = State) %>%
-  mutate(state_full = tolower(state_full)) %>%
-  left_join(us_state, by = "state_full" ) %>%
-  drop_na(`Crude Rate`)
-
-death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
-
-treatment_locations <- read_csv("Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
-  select(`Program Name`, Street, City, State, Zipcode)
+treatment_locations <- read_rds("treatment_locations.rds")
 
 # state populations
-
-state_pop <- read_csv("State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
-  rename(state_full = NAME) %>%
-  mutate(state_full = tolower(state_full))
-
+#
+# state_pop <- read_rds("state_pop.rds")
+#
+# us_state <- map_data("state") %>%
+#   rename(state_full = region)
+#
 # https://simplemaps.com/data/us-cities
 
-longlatinfo <- read_csv("uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
-  rename(City = city) %>%
-  rename(State = state_id)
+longlatinfo <- read_rds("longlatinfo.rds")
 
-treatment_locations_map <- treatment_locations %>%
-  left_join(longlatinfo, by = c("City", "State")) %>%
-  group_by(State) %>%
-  mutate(count = n()) %>%
-  mutate(state_name = tolower(state_name)) %>%
-  rename(state_full = state_name) %>%
[…]

### 2020-04-27 Opioid_Crisis_Code.Rmd
renamed

--- 2020-04-26 ms_5.Rmd
@@
        subtitle = "Between the Years of 2000 and 2014") +
-  geom_point() +
-  transition_reveal(Year)
+  geom_point()
 
@@
   clean_names() %>%
-  na.omit() %>%
+  filter(number_of_deaths != "N/A") %>%
   filter(age_range != "Total") %>%

### 2020-04-27 all_data.R
commented

@@ -14,6 +14,8 @@ library(plotly)
 library(DT)
 library(broom)
 
+# copied all data and turned into rds format to decrease size
+
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
 death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 

### 2020-04-27 Opioid_Crisis_App.R
README Fixed and GitHub link changed

@@ -265,7 +265,7 @@ tabPanel("Model",
                                    chapter55.digital.mass.gov."),
                          p("My project code can be found on my",
                            a("GitHub",
-                             href = "https://github.com/suruchiramanujan/final-project.git",)),
+                             href = "https://github.com/suruchiramanujan/Opioid-Crisis-Explore",)),
                          h1("About Me"),
                          p("My name is Suruchi Ramanujan and I am a 
                                    senior in Quincy House studying 

### 2020-04-27 Opioid_Crisis_App.R
commented through

@@ -16,20 +16,22 @@ library(broom)
 
 # Read in Data
 
-# https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
+# Data from Wonder CDC
 
 death_by_year_and_state <- read_rds("death_by_year_and_state.rds")
 
+# Read in treatment locations
+
 treatment_locations <- read_rds("treatment_locations.rds")
 
-# state populations
-#
+# This commented out data was too large to fit into the published app
+
 # state_pop <- read_rds("state_pop.rds")
-#
+
 # us_state <- map_data("state") %>%
 #   rename(state_full = region)
-#
-# https://simplemaps.com/data/us-cities
+
+# Data from https://simplemaps.com/data/us-cities used later in map data
 
 longlatinfo <- read_rds("longlatinfo.rds")
 
@@ -40,25 +42,39 @@ longlatinfo <- read_rds("longlatinfo.rds")
 #   mutate(state_name = tolower(state_name)) %>%
 #   rename(state_full = state_name) %>%
 #   left_join(us_state, by = "state_full")
-#
+
 # treatment_locations_map_per_cap <- treatment_locations_map %>%
 #   left_join(state_pop, by = "state_full") %>%
 #   group_by(State) %>%
 #   mutate(count_per_pop = n()/POPESTIMATE2019)
-# 
+
+# Joining all of the longitude latitude info to treatment locations and looking
+# at just Massachusetts.
+
 treatment_locations_map_mas <- treatment_locations %>%
   left_join(longlatinfo, by = c("City", "State")) %>%
   filter(State == "MA") %>%
   drop_na()
 
+# Read in county info
+
 counties <- read_rds("counties.rds")
+
+# Looking at the deaths per county in Massachusetts
 
 madeathbycounty <- read_rds("madeathbycounty.rds")
 
+# Looking at county population for per capita calcs
+
 countypop <- read_rds("countypop.rds")
+
+# Join opioid deaths by county w county info
 
 madeathbycountywlonglat <- madeathbycounty %>%
   left_join(counties, by = "Municipality")
+
+# Cleaning up the data to look only at the info we want and look at deaths per
+# county/time period
 
 madeathbycountywlonglat <- madeathbycountywlonglat %>%
   select(County,
[…]

### 2020-04-27 Opioid_Crisis_Code.Rmd
code w comments

@@
 ---
-title: "ms_5"
+title: "opioid_crisis_code"
 author: "Suruchi Ramanujan"
-date: "3/20/2020"
+date: "4/27/2020"
 output: html_document
@@
 
-```{r}
+```{r commondiseases, echo = FALSE}
 
-# total deaths per year in u.s. by cancer, drug overdose, cvd
+# total deaths per year in u.s. by cancer, drug overdose, cvd compared to 1999
 
-deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", col_types = "ldddddd") %>%
-  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
-  mutate(`Proportion of 1999 Deaths by Cancer` = `Deaths by Cancer`/549829) %>%
-  mutate(`Proportion of 1999 Deaths by Drugs` = `Deaths by Drug Overdose`/19122) %>%
-  mutate(`Proportion of 1999 Deaths by CVD` = `Deaths by CVD`/949900)
+deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", 
+                          col_types = "ldddddd") %>%
+  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, 
+         `Deaths by CVD`) %>%
+  mutate(`Proportion of 1999 Deaths by Cancer` = 
+           `Deaths by Cancer`/549829) %>%
+  mutate(`Proportion of 1999 Deaths by Drugs` = 
+           `Deaths by Drug Overdose`/19122) %>%
+  mutate(`Proportion of 1999 Deaths by CVD` = 
+           `Deaths by CVD`/949900)
+
+# this is to look at the total levels of deaths based on disease; pivotted
+# longer for ggplot
 
 totaldeathsbycommondiseases <- deathsbycommondiseases %>%
-  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
+  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, 
+         `Deaths by CVD`) %>%
   pivot_longer(.,
@@
   rename(`Cause of Death` = name)
+
+# just the info needed to map compared to 1999 proportions
 
@@
 
-```{r}
-plot_deathsbycommondiseases <- ggplot(totaldeathsbycommondiseases, aes(x = Year, y = Deaths, color = `Cause of Death`)) +
+```{r ggplot commondiseases, echo = FALSE}
+
+# plot year vs deaths for different diseases. transitions no longer work w
+# plotly
+
+plot_deathsbycommondiseases <- ggplot(totaldeathsbycommondiseases, aes(x = Year, 
+                                        y = Deaths, color = `Cause of Death`)) +
   geom_line() +
@@
 
-prop_deathsbycommondiseases <- ggplot(propdeathsbycommondiseases, aes(x = Year, y = Prop.Deaths, color = `Cause of Death`)) +
+# plot year vs proportion of 1999 deaths for different diseases. transitions no
+# longer work w plotly
+
+prop_deathsbycommondiseases <- ggplot(propdeathsbycommondiseases, 
+                    aes(x = Year, y = Prop.Deaths, color = `Cause of Death`)) +
   geom_line() +
@@
 
-```{r}
+```{r overdose_counts, echo = FALSE}
 
-# https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
+# data from wonder from cdc..got rid of unreliable data
 
-death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
-                                    col_types = "lcdccddddc") %>%
+death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv"
+, col_types = "lcdccddddc") %>%
   filter(`Crude Rate` != "Unreliable")
+
+# state info for map
 
@@
   rename(state_full = region)
+
+# only want death by drugs not alcohol or unknown. join w state info
 
@@
 
-death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)
+# change cruderate for plot
+
+death_by_year_and_state$`Crude Rate` <-
+  as.numeric(death_by_year_and_state$`Crude Rate`)
+
+# make a map of opioid death rates by state
 
@@
 
-```{r, message=FALSE}
+```{r notinuse, echo = FALSE}
+
 #drug treatment centers
 
-treatment_locations <- read_csv("Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
+treatment_locations <- read_csv("Drug Treatment Centers.csv", 
+                                col_types = "ccccccccc") %>%
   select(`Program Name`, Street, City, State, Zipcode)
@@
 
-# https://simplemaps.com/data/us-cities
+# https://simplemaps.com/data/us-cities longlatinfo
 
-longlatinfo <- read_csv("uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
+longlatinfo <- read_csv("uscitieslonglat.csv", 
+                        col_types = "ccccdcccddddcllcdcd") %>%
   rename(City = city) %>%
   rename(State = state_id)
+
+# treatment location info
 
@@
 
-# not in final
-
-ggplot(data = treatment_locations_map,
-            mapping = aes(x = long, y = lat.y,
-                          fill = count, group = group)) + 
-  geom_polygon(color = "gray90", size = 0.05) + 
-  theme_map() +
-  scale_fill_gradient(low = "white", high = "#CB454A",
-                      breaks = c(0,20,40,60,80,120)) +
-    guides(fill = guide_legend(nrow = 1)) + 
-    theme(legend.position = "bottom") +
-  labs(title = "Distribution of Opioid Treatment Centers in 2020",
-       color = "Number of Treatment Centers in State")
+# modify data to be per capita
 
@@
 
-```{r}
-# in final
+```{r treatmentcenters, echo = FALSE}
+
+# not in final but map of distribution of treatment centers per capita
 
@@
 
-```{r, message = FALSE}
-#MA specific data
+```{r madeathrates, echo = FALSE}
 
-madeathbyagegender <- read_csv("MAFatalOpioidOverdosesByAgeAndGender.csv",
-                               col_types = "ccddd")
+# read in opioid related death rate per year of ma vs that of us
 
-macountyheroin <- read_csv("MAHeroinAsPrimarySubstanceOfUse.csv",
-                           col_types = "ccccc")
+mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", 
+                         col_types = "cdc")
 
-madeathbyrace <- read_csv("MAOpioidDeathsByRaceAndEthnicity.csv",
-                          col_types = "dcdd")
-
-matypeofsubstance <- read_csv("MAPrimarySubstanceofUseWhenEnteringTreatmentbyTown.csv",
-                              col_types = "cccccccccccc")
-
-toxdeath <- read_csv("MAProportionDecedentsWithRxHistorybyToxicologyScreen.csv",
-         skip = 1,
-         col_types = "cdccccccdc")
-
-```
-
-```{r}
-mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", col_types = "cdc")
+# only looking at data for which there is complete info
 
@@
   filter (Year != "1999") %>%
-  rename(deathperhundredthousand =`Age-Adjusted Opioid-Related Death Rate per 100,000 People`)
+  rename(deathperhundredthousand =
+           `Age-Adjusted Opioid-Related Death Rate per 100,000 People`)
 
-plot_mavsus_death <- ggplot(mavsus_death, aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
+# ggplot grouped by geography
+
+plot_mavsus_death <- ggplot(mavsus_death, 
+      aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
   geom_line() +
@@
 
-```{r}
+```{r plot_1_ma, echo = FALSE}
+
+# county data
+
 counties <- read_csv("zip_codes_states.csv",
@@
 
-madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv",
+# opioid related deaths by county
+
+madeathbycounty <- read_csv(
+  "MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv",
                             col_types = "cdcdddc")
+
+# county population for per capita info
 
@@
 
+# counties + longitude/latitude
+
 madeathbycountywlonglat <- madeathbycounty %>%
   left_join(counties, by = "Municipality")
+
+# modifying and dividing by time and county/subregion
 
@@
   group_by(subregion) %>%
-  mutate(total_deaths_2001.5 = sum(`Confirmed Opioid Related Death Count 2001-2005`)) %>%
-  mutate(total_deaths_2006.10 = sum(`Confirmed Opioid Related Death Count 2006-2010`)) %>%
-  mutate(total_deaths_2011.15 = sum(`Confirmed Opioid Related Death Count 2011-2015`)) %>%
+  mutate(total_deaths_2001.5 = 
+          sum(`Confirmed Opioid Related Death Count 2001-2005`)) %>%
+  mutate(total_deaths_2006.10 = 
+          sum(`Confirmed Opioid Related Death Count 2006-2010`)) %>%
+  mutate(total_deaths_2011.15 = 
+           sum(`Confirmed Opioid Related Death Count 2011-2015`)) %>%
   distinct(subregion, .keep_all= TRUE)
+
+# county map info
 
@@
 
-# https://www.indexmundi.com/facts/united-states/quick-facts/massachusetts/percent-of-people-of-all-ages-in-poverty#table
+# poverty by county (not in use as of now)
 
@@
   mutate(subregion = tolower(subregion))
+
+# combining all mass data together
 
@@
 
+# 2001-2005 poverty percent vs opioid death rate per capita by county
+
 ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2001.5)) +
@@
 
+# not in use but correlation between poverty rate and opioid death per capita
+
 full_data %>%
   summarize(correlation = cor(percap_2001.5,`Poverty Percent`))
+
+# 2006-2010 poverty percent vs opioid death rate per capita by county
 
@@
 
+# not in use but correlation between poverty rate and opioid death per capita
+
 full_data %>%
   summarize(correlation = cor(percap_2006.10,`Poverty Percent`))
+
+# 2011-2015 poverty percent vs opioid death rate per capita by county
 
@@
 
+# not in use but correlation between poverty rate and opioid death per capita
+
 full_data %>%
   summarize(correlation = cor(percap_2011.15,`Poverty Percent`))
+
+# poverty percent by county (not in use)
 
@@
 
+# ggplot of opioid death rate per capita by county between 01 and 05
+
 ggplot(data = full_data,
@@
     theme(legend.position = "bottom")
+
+# ggplot of opioid death rate per capita by county between 06 and 10
 
@@
 
+# ggplot of opioid death rate per capita by county between 11 and 15
+
 ggplot(data = full_data,
@@
 
-# don't put these in final app
+# don't put these in final app bc they are the same graphs as above except w
+# total death info rather than per capita- will prob confuse reader
 
@@
 
-```{r}
+```{r leaflet, echo = FALSE}
 
-#treatment centers in MA
+# just looking at treatment centers in MA
 
@@
   drop_na()
+
+#treatment centers in MA distribution using leaflet
 
@@
 
-# need to figure this out bc color isnt working
- # addLegend(position = "bottomright", colors = "#0000FF", title = "Opioid Treatment Sites")
-
 ```
 
-```{r}
+```{r ageandrace, echo = FALSE}
 
-# models generated from data from kff.org
+# models generated from data from kff.org which took data from cdc
 
-# ageopioid
+# age vs opioid deaths over time
 
-ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", col_types = "dcc") %>%
+ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", 
+                           col_types = "dcc") %>%
   clean_names() %>%
@@
 
-raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
+# race vs opioid deaths over time
+
+raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", 
+                            col_types = "dcc") %>%
   clean_names() %>%
@@
   mutate(opioid_deaths = as.numeric(opioid_deaths))
+
+# combine the two for ggplot later
 
@@
 
+# ggplot for age over time vs opioid deaths
+
 ggplot(ageandracemodel, aes(year, deathsbyage, color = age_range)) +
@@
   geom_smooth(method = "lm", se = FALSE) +
-  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Age Range")
+  labs(x = "Year", y = "Number of Opioid Deaths in America", 
+       color = "Age Range")
+
+# ggplot for race over time vs opioid deaths
 
@@
 
-```{r}
+```{r race_lm, echo = FALSE}
+
 # filtering data by race to show effects of time on deaths in each race group
@@
   filter(race == "White, Non-Hispanic")
+
+# all simple linear regression model
 
@@
     rename(`Lower End` = conf.low) %>%
-    rename(`Upper End` = conf.high) %>%
-    gt()
+    rename(`Upper End` = conf.high)
 
@@
 
-```{r}
+```{r age_lm, echo = FALSE}
 
@@
 
-# showing effects of being in each age group on number of deaths idt this makes sense
+# showing effects of being in each age group on number of deaths idt this makes
+# sense so not including it yet
 
-averageageopioid_model <- lm(number_of_deaths ~ age_range, data = ageopioidmodel)
+averageageopioid_model <- lm(number_of_deaths ~ age_range, 
+                             data = ageopioidmodel)
 averageageopioid_model %>%

### 2020-05-05 Opioid_Crisis_Code.Rmd
added image and changed treatment graph to png

@@
     theme(legend.position = "bottom") +
-  labs(title = "Distribution of Opioid Treatment Centers in 2020 (Per Capita)",
-       color = "Number of Treatment Centers in State Per Capita")
+  labs(color = "Number of Treatment Centers in State Per Capita")
 
+ggsave("treatment_centers_per_capita.png")
 ```

### 2020-05-05 Opioid_Crisis_App.R

@@ -218,7 +218,8 @@ ui <- navbarPage(theme = shinytheme("flatly"),
                             from_max =2018,
                             choices = levels(
                               as.factor(death_by_year_and_state$Year))),
-                  fixedRow(column(7, plotOutput("overdose_counts")),
+                  fixedRow(column(7, plotOutput("overdose_counts"),
+                                  inline = TRUE),
                 column(3, offset = 1,
                 p("We cannot assume that the distribution of opioid overdose 
                 deaths per capita differs is uniform. There is certainly 
@@ -228,11 +229,10 @@ ui <- navbarPage(theme = shinytheme("flatly"),
                 deaths per capita over the past 20 years, beginning with some of 
                 the lowest death rates in 1998 and rising to have the highest 
                 state death rate per capita in 2018.")))),
-                br(),
-                
                 h3("Distribution of Opioid Treatment Centers in 2020 
                    (Per Capita)"),
-                fixedRow(column(7,plotlyOutput("treatment_centers_per_capita")),
+                fixedRow(column(7,imageOutput("treatment_centers_per_capita"),
+                                inline = TRUE),
                 column(3, offset = 1,
                 p("Unfortunately, the distribution of treatment centers per 
                 capita does not match up with the number of opioid deaths per 
@@ -324,8 +324,15 @@ tabPanel("Model",
          mainPanel(
            DTOutput("coefrace"))),
          tabPanel("About",
+                  imageOutput("pills"),
+                  br(),
+                  br(),
+                  h2("Visualizing the Effects of Opioids Across the United States", align = "center"),
+                  h4(em("An Analysis of Opioid Death Distribution"), align = "center"),
+                  br(),
+                  div(),
                   column(9,
-                         h1("Background"),
+                         h3("Background"),
                          p("Opioids have been prescribed for pain management for 
                          many decades, although they are effective at treating 
                          acute, not chronic pain. In treating chronic pain, 
@@ -357,7 +364,7 @@ tabPanel("Model",
                          p("Finally, I sought to create a model between factors 
                          of interest such as age and ethnicity and increases in 
                            opioid deaths over the past twenty years."),
-                         h1("The Data"),
+                         h3("The Data"),
                          p("Data concerning the entire United States came from 
[…]

### 2020-05-07 Opioid_Crisis_Code.Rmd
moved data to a folder

@@
 
-deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", 
+deathsbycommondiseases <- read_csv("raw-data/deathsbycommondiseases.csv", 
                           col_types = "ldddddd") %>%
@@
 
-death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv"
+death_by_year_and_state <- read_csv("raw-data/Drug Overdose Deaths by Year and State.csv"
 , col_types = "lcdccddddc") %>%
@@
 
-treatment_locations <- read_csv("Drug Treatment Centers.csv", 
+treatment_locations <- read_csv("raw-data/Drug Treatment Centers.csv", 
                                 col_types = "ccccccccc") %>%
@@
 
-state_pop <- read_csv("State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
+state_pop <- read_csv("raw-data/State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
   rename(state_full = NAME) %>%
@@
 
-longlatinfo <- read_csv("uscitieslonglat.csv", 
+longlatinfo <- read_csv("raw-data/uscitieslonglat.csv", 
                         col_types = "ccccdcccddddcllcdcd") %>%
@@
 
-# not in final but map of distribution of treatment centers per capita
+# map of distribution of treatment centers per capita
 
@@
 
-mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", 
+mavsus_death <- read_csv("raw-data/MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", 
                          col_types = "cdc")
@@
 
-counties <- read_csv("zip_codes_states.csv",
+counties <- read_csv("raw-data/zip_codes_states.csv",
                      col_types = "dddccc") %>%
@@
 madeathbycounty <- read_csv(
-  "MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv",
+  "raw-data/MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv",
                             col_types = "cdcdddc")
@@
 
-countypop <- read_csv("countypop.csv",
+countypop <- read_csv("raw-data/countypop.csv",
                       col_types = "cdd") %>%
@@
 
-povertybycounty <- read_csv("PovertyByCounty.csv",
+povertybycounty <- read_csv("raw-data/PovertyByCounty.csv",
                             col_types = "cd") %>%
@@
 
-ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", 
+ageopioidmodel <- read_csv("raw-data/U.S. age vs. opioid deaths.csv", 
                            col_types = "dcc") %>%
@@
 
-raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", 
+raceopioidmodel <- read_csv("raw-data/U.S. race vs. opioid deaths.csv", 
                             col_types = "dcc") %>%

### 2020-05-07 Opioid_Crisis_App.R

@@ -212,7 +212,7 @@ ui <- navbarPage(theme = shinytheme("flatly"),
                                    seeing sharp increases in death by
                                    opioid overdose."))),
                                    fluidPage(
-                  titlePanel("Drug Overdoses Per Capita by State Over Time"),
+                  h3("Drug Overdoses Per Capita by State Over Time"),
                   sliderTextInput("Year", "Year", 
                             from_min = 1998,
                             from_max =2018,
@@ -245,7 +245,8 @@ ui <- navbarPage(theme = shinytheme("flatly"),
                 recent than the most recent opioid death data."))
                )),
 tabPanel("Massachusetts Opioid Data",
-         h1("Comparison of Death Rates between Massachusetts and the 
+         h1("Opioid Deaths in Massachusetts"),
+         h3("Comparison of Death Rates between Massachusetts and the 
             United States Between the Years of 2000 and 2014"),
              fixedRow(column(7,  plotlyOutput("madeathrates", height = "100%")),
                       column(3, offset = 1, p("The graph to the left, created 
@@ -255,7 +256,7 @@ tabPanel("Massachusetts Opioid Data",
                       Despite efforts at expanding healthcare (for example, 
                       through Romneycare), opioid deaths stay consistently 
                       higher than the national average"))),
-                h2("Opioid Deaths in Massachusetts by County over Time"),
+                h3("Opioid Deaths in Massachusetts by County over Time"),
          br(),
                 # create drop down for user selecting time period
                 
@@ -274,7 +275,7 @@ tabPanel("Massachusetts Opioid Data",
                 although this makes sense given the high population of Sussex. 
                 Thus, in order to better demonstrate the distribution of deaths, 
                 I chose to map the total deaths per capita by county."))),
-                h2("Distribution of Opioid Treatment Centers Across 
+                h3("Distribution of Opioid Treatment Centers Across 
                    Massachusetts"),
                 fixedRow(column(7, fluidPage(
                   leafletOutput("masstreat"))),
@@ -367,7 +368,11 @@ tabPanel("Model",
                          h3("The Data"),
                          p("Data concerning the entire United States came from 
                          the CDC's Wonder Archive. Massachusetts-specific data 
-                         came from chapter55.digital.mass.gov."),
+                         came from chapter55.digital.mass.gov. The image above 
+                           came from",
+                           a("here",
[…]

### 2020-05-07 all_data.R

@@ -18,7 +18,7 @@ library(broom)
 
 # https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E
 
-death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv", 
+death_by_year_and_state <- read_csv("raw-data/Drug Overdose Deaths by Year and State.csv", 
                                     col_types = "lcdccddddc") %>%
   filter(`Crude Rate` != "Unreliable")
 
@@ -36,14 +36,14 @@ death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude
 
 write_rds(death_by_year_and_state, "death_by_year_and_state.rds")
 
-treatment_locations <- read_csv("Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
+treatment_locations <- read_csv("raw-data/Drug Treatment Centers.csv", col_types = "ccccccccc") %>%
   select(`Program Name`, Street, City, State, Zipcode)
 
 write_rds(treatment_locations, "treatment_locations.rds")
 
 # state populations
 
-state_pop <- read_csv("State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
+state_pop <- read_csv("raw-data/State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
   rename(state_full = NAME) %>%
   mutate(state_full = tolower(state_full))
 
@@ -51,7 +51,7 @@ write_rds(state_pop, "state_pop.rds")
 
 # https://simplemaps.com/data/us-cities
 
-longlatinfo <- read_csv("uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
+longlatinfo <- read_csv("raw-data/uscitieslonglat.csv", col_types = "ccccdcccddddcllcdcd") %>%
   rename(City = city) %>%
   rename(State = state_id)
 
[…]

### 2020-05-08 Opioid_Crisis_Code.Rmd
fixed legends

@@
   theme_map() +
-  scale_fill_gradient(low = "white", high = "#CB454A") +
+  scale_fill_gradient(name = "Overdose Deaths Per Capita/100,000",
+                            low = "white", high = "#CB454A",
+                            breaks = c(0,5,10,15,20,25,30,35,40,45,50)) +
     guides(fill = guide_legend(nrow = 1)) + 
@@
 # map of distribution of treatment centers per capita
+treatment_locations_map_per_cap <- treatment_locations_map_per_cap %>%
+  rename("Number of Treatment Centers Per Capita" = count_per_pop)
 
@@
             mapping = aes(x = long, y = lat.y,
-                          fill = count_per_pop, group = group)) + 
+                          fill = `Number of Treatment Centers Per Capita`, group = group)) + 
   geom_polygon(color = "gray90", size = 0.05) + 
   theme_map() +
-  scale_fill_gradient(low = "white", high = "#CB454A",
-                      breaks = c(0,20,40,60,80,120)) +
+  scale_fill_gradient(low = "white", high = "#CB454A") +
     guides(fill = guide_legend(nrow = 1)) + 
     theme(legend.position = "bottom") +
-  labs(color = "Number of Treatment Centers in State Per Capita")
+  labs(color = "Number of Treatment Centers Per Capita")
 

### 2020-05-08 Opioid_Crisis_App.R

@@ -35,19 +35,6 @@ treatment_locations <- read_rds("treatment_locations.rds")
 
 longlatinfo <- read_rds("longlatinfo.rds")
 
-# treatment_locations_map <- treatment_locations %>%
-#   left_join(longlatinfo, by = c("City", "State")) %>%
-#   group_by(State) %>%
-#   mutate(count = n()) %>%
-#   mutate(state_name = tolower(state_name)) %>%
-#   rename(state_full = state_name) %>%
-#   left_join(us_state, by = "state_full")
-
-# treatment_locations_map_per_cap <- treatment_locations_map %>%
-#   left_join(state_pop, by = "state_full") %>%
-#   group_by(State) %>%
-#   mutate(count_per_pop = n()/POPESTIMATE2019)
-
 # Joining all of the longitude latitude info to treatment locations and looking
 # at just Massachusetts.
 
@@ -229,8 +216,13 @@ ui <- navbarPage(theme = shinytheme("flatly"),
                 deaths per capita over the past 20 years, beginning with some of 
                 the lowest death rates in 1998 and rising to have the highest 
                 state death rate per capita in 2018.")))),
-                h3("Distribution of Opioid Treatment Centers in 2020 
-                   (Per Capita)"),
+                br(),
+                br(),
+                br(),
+                br(),
+                br(),
+                br(),
+                h3("Distribution of Opioid Treatment Centers Per Capita in 2020"),
                 fixedRow(column(7,imageOutput("treatment_centers_per_capita"),
                                 inline = TRUE),
                 column(3, offset = 1,
@@ -271,8 +263,9 @@ tabPanel("Massachusetts Opioid Data",
                 distribution of opioid deaths per capita across the counties of 
                 Massachusetts for three different time periods between 2000 and 
                 2015. Having looked at the data for total opioid deaths 
-                (not per data), deaths seem to be concentrated in Sussex county, 
-                although this makes sense given the high population of Sussex. 
+                (not per capita), I noticed that deaths seem to be concentrated in 
+                Middlesex county, 
+                although this makes sense given the high population of Middlesex. 
                 Thus, in order to better demonstrate the distribution of deaths, 
                 I chose to map the total deaths per capita by county."))),
                 h3("Distribution of Opioid Treatment Centers Across 
@@ -391,7 +384,7 @@ tabPanel("Model",
 )))
 server <- function(input, output, session) {
   
[…]
